---
title: "Mental health trajectories of religious majority and minority youth"
format: 
  html: 
    toc: true
    toc-depth: 3
    number-sections: true
    df-print: paged
    embed-resources: true
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

This Quarto document pre-processes data from the Young Lives project, accessed via the UK Data Service on 4th August, 2026.

# Set up

*Load packages*

```{r setup}
#| code-fold: true

# Install any packages that aren't there yet then load
pacman::p_load(
  tidyverse, # data wrangling
  here, # clean file paths
  lavaan, # latent growth curve models
  haven, 
  patchwork, 
  purrr,
  ggplot2
)

# Set colours
group_cols <- c(Majority = "#4A90D9",   # blue
                Minority = "#E05B5B")   # red

options(scipen = 999)
```

*Load data*

```{r load-data}
#| code-fold: true

# here() builds a path from project root, so it works regardless of machine or subfolder
countries <- c("ethiopia", "india", "peru", "vietnam")

# Load rounds 1-7 constructed files 
data_list <- map(countries, ~ read_sav(here("data", paste0(.x, ".sav")))) |>
  set_names(countries)

# Load SDQ-E files (rounds 3-5, 1994-cohort only)
rounds    <- 3:5
combos <- tidyr::expand_grid(country = countries, round = rounds) # every country–round combination
sdq_list <- map2(
  combos$country, combos$round,
  ~ read_sav(here("data", paste0(.x, "_r", .y, "_sdq.sav")))
) |>
  set_names(paste0(combos$country, "_r", combos$round, "_sdq"))

# Note: Generally, the SDQ files for rounds 3-5 are named occh_older_child. The only exception is for Peru rounds 4 and 5, the files are named olderchild-SAQ. 
```

### Variable transformations

Key variables from each file

- Constructed files (ethiopia, india, peru, vietnam): child's ID (*childid*), cohort (*yc*), round, cluster ID (*clustid),* age in years (*ageyea*)*,* **sex** (*chsex*), child's ethnic group (*chethnic*), **child's religious affiliation** (*chldrel*), **life satisfaction** (*cladder*)

- SDQ-E files (ethiopia_r3-5, india_r3-5, peru_r3-5, vietnam_r3-5): strengths and difficulties questionnaire emotional subscale - **psychological distress** (comprised of 5 items) collected at rounds 3, 4 and 5 for the 1994-cohort at ages 15, 19, and 22 respectively

#### Select variables

*Constructed file*

```{r main-file-variables}
#| code-fold: true

# Variables from constructed file
keep_vars <- c("childid", "yc", "round", "ageyea", "chldrel", "cladder", "chsex", "chethnic", "clustid", "careladder")
data <- map(data_list, function(df) {
  df |> 
    select(all_of(keep_vars)) |> 
    mutate(childid = as.character(childid))
})
```

*SDQ-E files*

```{r sdq-variables}
#| code-fold: true

## Country prefixes for childid
prefixes <- c(ethiopia = "ET", india = "IN", peru = "PE", vietnam = "VN")

## Fix childid: make sure there are 6 digits and add country prefix
fix_childid <- function(id, prefix) {
  id <- as.character(id)
  ifelse(str_detect(id, "^[A-Za-z]"),                 # if already full format
         id,                                          # leave it alone
         paste0(prefix, str_pad(id, 6, "left", "0"))) # otherwise reconstruct
}

## SDQ items come in two naming schemes across countries/rounds
scheme_headac <- function(r) c(
  headache = paste0("HEADACR", r),
  worry    = paste0("WRYLOTR", r),
  unhappy  = paste0("UNHPPYR", r),
  nervous  = paste0("NRVSITR", r),
  fears    = paste0("MNYFERR", r)
)
scheme_feay <- function(r) c(
  headache = paste0("FEAY13R", r),
  worry    = paste0("FEAY05R", r),
  unhappy  = paste0("FEAY20R", r),
  nervous  = paste0("FEAY38R", r),
  fears    = paste0("FEAY28R", r)
)

## Which scheme each country uses at R4/R5 (R3 is same across countries)
country_scheme <- list(
  india    = scheme_feay,
  peru     = scheme_headac,
  ethiopia = scheme_feay,
  vietnam  = scheme_feay
)

## Clean SDQ files: select + rename items and ID, then fix childid
sdq <- imap(sdq_list, function(df, nm) {
  ctry <- str_extract(nm, "^[a-z]+")
  rnum <- str_extract(nm, "[3-5]")
  pfx  <- prefixes[[ctry]]
  
  df <- df |> rename_with(toupper)
  
  # R3 uses HEADAC everywhere; R4/R5 use the country's scheme
  items <- if (rnum == "3") scheme_headac(rnum) else country_scheme[[ctry]](rnum)
  items <- setNames(items, paste0(names(items), "_r", rnum))
  
  idc <- if (rnum == "3") "CHILDID" else "CHILDCODE"
  
  df |>
    select(all_of(c(childid = idc, items))) |>
    mutate(childid = fix_childid(childid, pfx))
})
```

#### Check SDQ

*Check values*

```{r inspect-sdq}
#| code-fold: true

# Remove missing data 
missing <- c(77,79,88,99)
sdq <- imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  df |> 
    mutate(across(all_of(cols),
                  ~ if_else(as.numeric(.x) %in% missing, NA_real_,
                            as.numeric(.x))))
})

## Distinct response values in each country-round
imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  sort(unique(as.numeric(unlist(df[cols]))))
})
```

There are different response options depending on rounds - for India, Ethiopia, and Vietnam, round 3 has 3 response options but rounds 4-5 have 4. For Peru, there are 3 response options across time.

#### Transform variables

*Check values of variables - life satisfaction, sex, religion, ethnicity/caste, age*

```{r check-variables}
#| include: false

# Check distinct values of each key variable 
map(data, function(df) {
  map(df[c("cladder", "chsex", "chldrel", "chethnic", "ageyea", "careladder")], ~sort(unique(as.numeric(.x))))
})
# No NA values 77/88/99 to transform

```

```{r variables-named}
#| code-fold: true 

# Transform into named categories
map(data, ~ haven::as_factor(.x$chldrel) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chsex) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chethnic) |> table(useNA = "always"))
```

*Create separate dataframes for each country*

```{r country-dataframes}
#| code-fold: true

sdq_items <- c("headache", "worry", "unhappy", "nervous", "fears")
# SDQ: keep individual items as well as total
sdq_scores <- imap(sdq, function(df, nm) {
  rnum  <- str_extract(nm, "[3-5]")
  items <- paste0(sdq_items, "_r", rnum)          # headache_r3, ... fears_r3
  total <- paste0("emo_r", rnum)                  # keep the total too 
  keep  <- intersect(c(items, total), names(df))  # whatever's present
  df |> select(childid, all_of(keep))
})

# Merge constructed data and SDQ for each country 
countries <- c("ethiopia", "india", "peru", "vietnam")
merged_data <- map(countries, function(ctry) {
  base      <- data[[ctry]]
  sdq_files <- sdq_scores[str_starts(names(sdq_scores), paste0(ctry, "_"))]
  reduce(sdq_files, left_join, by = "childid", .init = base)
}) |>
  set_names(countries)

# Row counts shouldn't jump (a jump = duplicate childids somewhere)
map_int(merged_data, nrow)

# Create separate dataframes for each country
list2env(merged_data, envir = .GlobalEnv) 

# For each child, take round 1's clustid for clustering
r1_clust <- map(merged_data, ~ .x |> filter(round ==1) |> 
                  mutate(clustid = if_else(clustid %in% c(0,80,88,90,99), NA_real_, as.numeric(clustid))) |> 
                  select(childid, clustid_base = clustid))
```

### Structuring data

#### Create variable 'equivalent age'

Although we have age in years as the variable for real age, we need an equivalent age of each round i.e. in round 3 for example, the 2001-cohort is aged 8 and the 1994-cohort is aged 15.

```{r equivalent-age}
#| code-fold: true

age_yc <- c("1"=1, "2"=5, "3"=8, "4"=12, "5"=15, "6"=19, "7"=22)
age_oc <- c("1"=8, "2"=12, "3"=15, "4"=19, "5"=22, "6"=26, "7"=29)

add_age <- function(df) {
  df %>%
    mutate(equivalent_age = if_else(yc == 1, age_yc[as.character(round)],
                                    age_oc[as.character(round)])) %>%
    filter(!is.na(equivalent_age)) %>%
    group_by(childid, equivalent_age) %>%
    slice_max(!is.na(cladder), n = 1, with_ties = FALSE) %>% # In round 6 there are multiple rows for phone calls but we only keep the row that has cladder data
    ungroup()
}

stacked <- map(merged_data, add_age)

stacked <- map(stacked, \(df) {
  df |> mutate(
    # pick the item set for whichever round this row belongs to
    sdq = dplyr::case_when(
      round == 3 ~ rowSums(across(all_of(paste0(sdq_items, "_r3"))), na.rm = FALSE),
      round == 4 ~ rowSums(across(all_of(paste0(sdq_items, "_r4"))), na.rm = FALSE),
      round == 5 ~ rowSums(across(all_of(paste0(sdq_items, "_r5"))), na.rm = FALSE),
      TRUE ~ NA_real_
    )
  )
})
names(stacked$ethiopia)   # emo_r3-5 combined into sdq
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Social determinants categorisations

::: callout-note
**We include the following social determinants:**

Sex, religion, race/ethnic/caste, and socioeconomic status (TBC)
:::

Sex is very clearly two categories in the dataset already, however we have to specify which religions and ethnic groups/castes are majority/minority.

Rather than using the group sizes in this sample, which are not representative of respective countries, we take into account country contexts for categorisations.

## Categorise religious minority and majority status

We take into account country context when classifying majority/minority rather than solely rely on World Religion Database.

World Religion Database religious demographics from 2000:

- Ethiopia: 58% Christian, 34% Muslim, 7.0% Ethnic religionists

- India: 75% Hindus, 13% Muslims, 4.2% Christians, 3.3% Ethnic religionists, 1.8% Sikhs, 1.4% Agnostics

- Peru: 96% Christian, 1.4% Ethnic religionist, 1.2% Agnostic

- Vietnam: 48% Buddhist, 13% Agnostic, 12% Ethnic religionist, 10.8% New religionist, 8.3% Christian, 6.7% Atheist, 1.0% Chinese folk religionist

*Check sample n by religion*

```{r check-religions}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chldrel) |>
      count(chldrel = as_factor(chldrel)))
```

For now, we will define the majority religions for each country as the following

- Ethiopia: **Orthodox**
- India: **Hinduism**
- Peru**: Catholicism**
- Vietnam: **Buddhism, none, and ancestor worship** (to be discussed more, but all of these could be considered dominant traditions in Vietnam)

*Classify majority/minority*

```{r classify}

# Specify majority religions
majority_religion <- list(ethiopia = "orthodox", india = "hindu", peru = "catholic", vietnam = c("buddhist", "none", "ancestor worship"))

classify <- function(df, ctry) {
  maj <- majority_religion[[tolower(ctry)]]
  df |> 
    mutate(
      chldrel = as.character(as_factor(chldrel)),
     religious_minority = case_when(
       is.na(chldrel) ~ NA_character_,
       tolower(chldrel) %in% maj ~ "Majority",
       TRUE ~ "Minority"
     )
      )
}

stacked <- imap(stacked, classify)
```

As sensitivity check in later analyses, if there are variations across Young Lives regions, instead of country-level demographics we will use regional demographics to determine majority and minority status.

*Check classifications*

```{r check-classifications}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chldrel, religious_minority) |>
      count(chldrel, religious_minority))

map(stacked, ~ .x |>
      distinct(childid, chldrel, religious_minority) |>
      count(religious_minority) |>
      mutate(pct = round(100 * n / sum(n), 1)))

# Check how many NAs, then drop
stacked <- imap(stacked, classify) |> map(~ filter(.x, !is.na(religious_minority)))
```

Missingness for religion (*chldrel*)

- Ethiopia: 2

- India: 0

- Peru: 1

- Vietnam: 1

------------------------------------------------------------------------

------------------------------------------------------------------------

## Categorise ethnic/caste minority and majority status

**As another social determinant we add race/ethnicity and caste. Caste is the same variable as race for India.**

In some countries e.g. Ethiopia it is difficult to categorise majority vs minority status, but Oromo (34-35%) and Amhara (27%) together represent more than half of the country's population, according to the 2007 census.

*Check sample n by race / caste*

```{r check-race}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chethnic) |>
      count(chethnic = as_factor(chethnic)))
```

For now, we will define the majority ethnic/racial groups or caste for each country as the following

- Ethiopia: **Amhara, Oromo** (34-35%, 27% - together more than half)
- India: **Other, Hindu** (scheduled castes SC and scheduled tribes ST most disadvantaged)
- Peru**: Mestizo, White** (Mestizo 60%, white Peruvians only 4-5% of population but a lot of power)
- Vietnam: **Kinh** (Viet) (about 85-86% of population)

*Classify ethnic/racial majority/minority*

```{r classify-race-caste}

# Specify majority race/ethnicity/caste
maj_ethnic <- list(ethiopia = c("Amhara", "Oromo"), india = "Other, Hindu", peru = c("Mestizo", "White"), vietnam = "Kinh")

classify_ethnic <- function(df, ctry) {
  maj <- maj_ethnic[[tolower(ctry)]]
  df |> 
    mutate(
      chethnic = as.character(as_factor(chethnic)),
     ethnic_minority = case_when(
       is.na(chethnic) ~ NA_character_,
       chethnic %in% maj ~ "Majority",
       TRUE ~ "Minority"
     )
      )
}

stacked <- imap(stacked, classify_ethnic)
```

As sensitivity check in later analyses, if there are variations across Young Lives regions, instead of country-level demographics we will use regional demographics to determine majority and minority status.

*Check classifications*

```{r check-classifications-ethnic}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chethnic, ethnic_minority) |>
      count(chethnic, ethnic_minority))

map(stacked, ~ .x |>
      distinct(childid, chethnic, ethnic_minority) |>
      count(ethnic_minority) |>
      mutate(pct = round(100 * n / sum(n), 1)))

# Check how many NAs, then drop
stacked <- imap(stacked, classify) |> map(~ filter(.x, !is.na(ethnic_minority)))
```

Missingness for race/ethnicity/caste (*chethnic*): Ethiopia n=2

------------------------------------------------------------------------

------------------------------------------------------------------------

# Descriptives

## Visualisations

All visualisations will be for the 1994-cohort only (unless comparing cohorts), in line with analyses.

2001-cohort will be used as a replication sample for life satisfaction models.

### Life satisfaction

*Overall trajectory*

```{r overall-traj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory <- function(df, ctry) {
  # split version
  df |>
    filter(yc == 0) |> 
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(2, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory)
wrap_plots(plots, ncol = 2)
```

*By cohort*

```{r traj-cohort}
#| code-fold: true

# Mean cladder by age and cohort, at overlap ages, per country
cohort_check <- map(stacked, ~ .x |>
                      filter(!is.na(cladder), equivalent_age %in% c(8, 12, 15, 19, 22, 26, 29)) |>
                      mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
                      group_by(equivalent_age, cohort) |>
                      summarise(mean_cl = mean(cladder), se = sd(cladder)/sqrt(n()), n = n(),
                                .groups = "drop"))

cohort_check$ethiopia

# Plot 
bind_rows(stacked, .id = "country") |>
  filter(!is.na(cladder), equivalent_age %in% c(8, 12, 15, 19, 22, 26, 29)) |>
  mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
  group_by(country, equivalent_age, cohort) |>
  summarise(mean_cl = mean(cladder), se = sd(cladder)/sqrt(n()), .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = cohort, fill = cohort)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  facet_wrap(~ country) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  labs(y = "Mean life satisfaction", x = "Age") +
  theme_minimal()
```

***By sex (1994-cohort only)***

```{r traj-sex}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_sex <- function(df, ctry) {
  # split version
  df |>
    filter(yc == 0) |> 
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    mutate(chsex = as_factor(chsex)) |> 
    group_by(equivalent_age, chsex) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = chsex)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(3, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sex)
wrap_plots(plots, ncol = 2)
```

***By religious minority/majority status (1994-cohort only)***

```{r traj-min-maj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_minmaj <- function(df, ctry) {
  # split version
  df |>
    filter(yc == 0) |> 
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age, religious_minority) |>
    summarise(mean_cl = mean(cladder),
              se = sd(cladder)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = religious_minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(2, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = religious_minority), alpha = 0.2, colour = NA) +
    scale_colour_manual(values = group_cols) +
    scale_fill_manual(values = group_cols) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_minmaj)
wrap_plots(plots, ncol = 2)
```

***By ethnic/racial/caste majority status (1994-cohort only)***

```{r traj-ethnic}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_ethnic_minmaj <- function(df, ctry) {
  # split version
  df |>
    filter(yc == 0) |> 
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age, ethnic_minority) |>
    summarise(mean_cl = mean(cladder),
              se = sd(cladder)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = ethnic_minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(2, 8)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = ethnic_minority), alpha = 0.2, colour = NA) +
    scale_colour_manual(values = group_cols) +
    scale_fill_manual(values = group_cols) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_ethnic_minmaj)
wrap_plots(plots, ncol = 2)
```

*By religion and sex (1994-cohort only)*

```{r traj-sex-relminority}
#| code-fold: true

bind_rows(stacked, .id = "country") |>
  filter(yc == 0) |> 
  filter(!is.na(cladder), !is.na(equivalent_age)) |>
  mutate(chsex = as_factor(chsex)) |>
  group_by(country, equivalent_age, chsex, religious_minority) |>
  summarise(mean_cl = mean(cladder),
            se = sd(cladder) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = religious_minority, fill = religious_minority)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96 * se, ymax = mean_cl + 1.96 * se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  scale_colour_manual(values = group_cols) +
  scale_fill_manual(values = group_cols) +
  facet_grid(country ~ chsex) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  scale_y_continuous(limits = c(2, 7)) +
  theme_minimal()
```

*By race/ethnicity/caste and sex (1994-cohort only)*

```{r traj-sex-ethnicminority}
#| code-fold: true

bind_rows(stacked, .id = "country") |>
  filter(yc == 0) |> 
  filter(!is.na(cladder), !is.na(equivalent_age)) |>
  mutate(chsex = as_factor(chsex)) |>
  group_by(country, equivalent_age, chsex, ethnic_minority) |>
  summarise(mean_cl = mean(cladder),
            se = sd(cladder) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = ethnic_minority, fill = ethnic_minority)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96 * se, ymax = mean_cl + 1.96 * se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  scale_colour_manual(values = group_cols) +
  scale_fill_manual(values = group_cols) +
  facet_grid(country ~ chsex) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  scale_y_continuous(limits = c(2, 7)) +
  theme_minimal()
```

------------------------------------------------------------------------

### Psychological distress (SDQ-E)

We cannot visualise SDQ properly because the measure is not invariant across time so descriptive visualisations would not be comparable.

We can only do per-item plots, for each round across countries. In Peru the same scale is used across countries so it can be plotted.

```{r sdq-country-rel}
#| code-fold: true

# Long format: one row per person-item-wave, with the raw response category
item_long <- imap_dfr(stacked, function(df, ctry) {
  df |>
    filter(yc == 0) |>
    select(childid, any_of(as.vector(outer(sdq_items, c("r3","r4","r5"),
                                            \(a,b) paste0(a,"_",b))))) |>
    pivot_longer(-childid, names_to = "item_wave", values_to = "response") |>
    filter(!is.na(response)) |>
    mutate(country = ctry,
           item = str_remove(item_wave, "_r[3-5]$"),
           wave = dplyr::recode(str_extract(item_wave, "r[3-5]$"),
                                r3 = "Age 15", r4 = "Age 19", r5 = "Age 22"))
})

# Proportion in each response category, per country x item x wave
item_props <- item_long |>
  count(country, item, wave, response) |>
  group_by(country, item, wave) |>
  mutate(prop = n / sum(n)) |>
  ungroup()

# Plot one item at a time (e.g. worry), facet by wave, countries side by side
item_props |>
  filter(item == "worry") |>
  ggplot(aes(x = factor(response), y = prop, fill = wave)) +
  geom_col(position = "dodge") +
  facet_wrap(~ country) +
  labs(title = "SDQ item: worry — response distribution across waves, by country",
       x = "Response category", y = "Proportion", fill = "Wave") +
  scale_fill_manual(values = c("Age 15" = "#4477AA",
                               "Age 19" = "#CC6677",
                               "Age 22" = "#DDCC77")) +
  theme_minimal()

# Peru only - check for religious maj/min
item_long <- imap_dfr(stacked, function(df, ctry) {
  df |>
    filter(yc == 0, !is.na(religious_minority)) |>
    mutate(group = if_else(religious_minority == "Minority" | religious_minority == 1, "Minority", "Majority")) |>
    select(childid, group,
           any_of(as.vector(outer(sdq_items, c("r3","r4","r5"), \(a,b) paste0(a,"_",b))))) |>
    pivot_longer(-c(childid, group), names_to = "item_wave", values_to = "response") |>
    filter(!is.na(response)) |>
    mutate(country = ctry,
           item = str_remove(item_wave, "_r[3-5]$"),
           wave = dplyr::recode(str_extract(item_wave, "r[3-5]$"),
                                r3="15", r4="19", r5="22"))
})
```

```{r sdq-maj-min}
#| code-fold: true

# Check size of smallest response category by religious majority/minority status - are some categories very sparse?
item_long |>
  count(country, wave, item, response) |>          
  group_by(country, wave, item) |>
  summarise(min_cell = min(n), .groups = "drop") |>
  arrange(min_cell) |> head(20)

# Raw counts per response category, split by group — look for tiny cells
sparse_check <- item_long |>
  count(country, group, wave, item, response) |>
  tidyr::pivot_wider(names_from = response, values_from = n, values_fill = 0)

# Flag the worry item as an example; scan all
sparse_check |> dplyr::filter(item == "worry") |> print(n = Inf)

# Quickest summary: smallest cell per country x group x wave x item
item_long |>
  count(country, group, wave, item, response) |>
  group_by(country, group, wave, item) |>
  summarise(min_cell = min(n), .groups = "drop") |>
  arrange(min_cell) |>
  head(20)
```

Looks alright, even for the smallest bin there are 90 people, and enough of individuals from religious minority group.

------------------------------------------------------------------------

## Summary table

*Set up to run for all exposures*

```{r exposure-setup}
#| code-fold: true

exposures <- c("religious_minority", "ethnic_minority", "chsex") # add SES

# baseline level goes second, difference is computed as level1 - level2
exp_levels <- list(
  religious_minority = c("Minority", "Majority"),
  ethnic_minority = c("Minority", "Majority"),
  chsex = c("male", "female")
)

pal <- c("#E05B5B", "#4A90D9")
exp_cols <- map(exp_levels, ~ set_names(pal, .x))

stacked <- map(stacked, ~ mutate(.x, chsex = as_factor(chsex)))
all_df <- bind_rows(stacked, .id = "country")
```

*Summary table: per country x age x determinant*

```{r table-descriptives}
#| code-fold: true

# Pivot longer, mean/SD/n per country x age x minority group
describe_exp <- function(df, exp) {
  df |> 
    pivot_longer(c(cladder, sdq, careladder), names_to = "outcome", values_to = "value") |> 
    filter(!is.na(value), !is.na(.data[[exp]])) |> 
    group_by(country, outcome, equivalent_age, group = .data[[exp]]) |> 
    summarise(mean = mean(value), sd = sd(value), n = n(), .groups = "drop") |> 
    mutate(cell = sprintf("%.2f (%.2f) [n=%d]", mean, sd, n)) |> 
    select(country, outcome, equivalent_age, group, cell) |> 
    pivot_wider(names_from = group, values_from = cell) |> 
    arrange(country, outcome, equivalent_age)
}

desc_tables <- set_names(exposures) |> map(~ describe_exp(all_df, .x))
desc_tables$religious_minority
desc_tables$ethnic_minority
desc_tables$chsex
```

## T Test

```{r comparisons}
#| code-fold: true

compare_exp <- function(df, exp) {
  df |> 
  filter(yc == 0) |>  # 1994-cohort only
  pivot_longer(c(cladder, sdq), names_to = "outcome", values_to = "value") |> 
  filter(!is.na(value), !is.na(.data[[exp]])) |> 
  mutate(grp = factor(.data[[exp]], levels = exp_levels[[exp]])) |> 
  group_by(country, outcome, equivalent_age) |> 
  filter(n_distinct(grp) == 2,
         min(table(grp)) >= 2) |> 
  summarise(
    lvl1 = exp_levels[[exp]][1], lvl2 = exp_levels[[exp]][2],
    n1 = sum(grp == lvl1), n2 = sum(grp == lvl2),
    tt = list(t.test(value ~ grp)),
    .groups = "drop"
  ) |> 
  mutate(
    diff = map_dbl(tt, ~ .x$estimate[1] - .x$estimate[2]), # lvl1 - lvl2
    ci_low = map_dbl(tt, ~ .x$conf.int[1]),
    ci_high = map_dbl(tt, ~ .x$conf.int[2]),
    p_value = map_dbl(tt, "p.value")
  ) |> 
  select(-tt)
}

comparisons <- set_names(exposures) |> map(~ compare_exp(all_df, .x))
comparisons$religious_minority
comparisons$ethnic_minority
comparisons$chsex
```

Welch's t test is preferred since the minority and majority groups do not have equal variances nor sample sizes. We can test whether the group means between minorities and majorities are statistically significant.

**Note that these are purely descriptive, we're doing many independent t-tests with no correction so this is more for establishing overall patterns.**

*diff_min_maj* is the minority - majority, i.e. if *diff_min_maj* is positive for *cladder*, that means that minorities have higher life satisfaction; whereas if it's positive for *sdq*, that means minorities have lower psychological distress (since the two scales are coded differently).

**Religious minority**

- Ethiopia: minority better life satisfaction at age 19, 26, 29

- India: minority better life satisfaction at age 22

**Ethnic minority**

- Ethiopia: minority better life satisfaction at age 19, 26; higher psychological distress at age 15, but lower distress at age 22

- India: minority lower life satisfaction at age 12, 15, 19, 22, 26, 29 (all ages!); higher psychological distress at age 15, 19 \[*clearest*\]

- Peru: minority higher life satisfaction at age 15

- Vietnam: minority lower life satisfaction at age 12, 15, 19, 22; higher distress at age 19 \[*clearest*\]

**Sex**

- Ethiopia: boys higher distress than girls at age 15, but girls higher distress than boys at age 19, 22

- India: girls better life satisfaction than boys at age 15, 26, 29; but girls higher distress than boys at age 15, 19, 22 (all ages) - girls higher distress but also higher life satisfaction

- Peru: girls better life satisfaction than boys at age 12, 15, 26; boys also higher distress than girls at age 15, 19, 22 (all ages) - boys higher distress and lower life satisfaction

- Vietnam: girls better life satisfaction than boys at age 15, 26; but girls higher distress than boys at age 15, 19, 22 (all ages) - girls higher distress but also higher life satisfaction

Interesting results from sex, in both India and Vietnam, girls have higher distress at 15,19,22 but also higher life satisfaction at 15,26. In Peru, boys have higher distress at 15,19,22 and lower life satisfaction at 12,15,26.

------------------------------------------------------------------------

## Histograms 

```{r histograms-cladder}
#| code-fold: true

plot_hist <- function(df, ctry, outcome_var, exp) {
  df |> 
    mutate(grp = factor(.data[[exp]], levels = exp_levels[[exp]])) |> 
    filter(!is.na(.data[[outcome_var]]), !is.na(grp)) |> 
    ggplot(aes(x = .data[[outcome_var]], fill = grp)) +
    geom_histogram(aes(y = after_stat(density)),
                   binwidth = 1, alpha = 0.5, position = "identity") +
    facet_wrap(~ equivalent_age, ncol = 4) +
    labs(title = paste0(ctry, " - ", outcome_var, " by ", exp), x = outcome_var, y = "Density", fill = NULL) +
    theme_minimal() +
    scale_fill_manual(values = exp_cols[[exp]]) +
    theme(legend.position = "bottom")
}

walk(exposures, function(exp) {
  iwalk(stacked, ~ print(plot_hist(.x, .y, "cladder", exp)))
})
```

```{r histograms-sdq}
walk(exposures, function(exp) {
  iwalk(stacked, ~ print(plot_hist(.x, .y, "sdq", exp)))
})
```

```{r histograms-careladder}
walk(exposures, function(exp) {
  iwalk(stacked, ~ print(plot_hist(.x, .y, "careladder", exp)))
})
```

------------------------------------------------------------------------

# Latent Growth Curve Analyses (LGCA)

*Structure data*

```{r pivot-wide}
#| code-fold: true

sdq_items <- c("headache","worry","unhappy","nervous","fears")
item_cols <- as.vector(outer(sdq_items, c("r3","r4","r5"), \(a,b) paste0(a,"_",b)))

# level1 -> 1, level2 -> 0 (keeps sign level1 - level2, matching t-tests)
# i.e. minority - majority; male - female
exp_code <- function(x, exp) if_else(as.character(x) == exp_levels[[exp]][1], 1, 0)

# Complete cases on exposure + covariates 
covs_for <- function(exp) if (exp == "chsex") "chsex" else c(exp, "chsex")

to_wide <- function(df) {
  wide_scores <- df |>
    mutate(across(all_of(exposures), ~ exp_code(.x, cur_column()))) |> 
    select(childid, all_of(exposures), clustid, yc, equivalent_age, cladder, careladder) |>
    pivot_wider(names_from = equivalent_age,
                values_from = c(cladder, careladder),
                names_glue = "{.value}_age{equivalent_age}")
  wide_items <- df |>  
    select(childid, any_of(item_cols)) |> 
    group_by(childid) |> 
    summarise(across(any_of(item_cols), ~first(na.omit(.x))), .groups = "drop")
  left_join(wide_scores, wide_items, by = "childid")
}
wide <- map(stacked, to_wide)
wide <- map(wide, ~ .x |> group_by(childid) |>
              fill(clustid, .direction = "downup") |> ungroup())

names(wide$ethiopia)  # check column names
```

*Filter children* (only those with 3 complete cladder out of 5 for 2001-cohort, 7 for 1994-cohort)

```{r filter-sample}
#| code-fold: true

wide <- map(wide, ~ .x |> 
              mutate(n_cl = rowSums(!is.na(across(starts_with("cladder_age"))))) |> 
              filter(n_cl >= 3) |> 
              select(-n_cl))
```

*Sample size*

```{r sample-size-oc}
#| code-fold: true

# 1994-cohort (yc == 0), with non-missing minority and sex (analytic filter)
older_n <- imap_dfr(set_names(exposures), function(exp, nm) 
  imap_dfr(wide, function(df, ctry) {
  d <- df |> filter(yc == 0, if_all(all_of(covs_for(exp)), ~ !is.na(.x)))
  tibble(exposure = nm, country = ctry, n = nrow(d))
}))
older_n
```

## Life satisfaction models

Given the two cohorts are not equivalent - with diverging trajectories, we will fit the models separately, and religious minority status and sex predicting the growth factors. 2001-cohort will have age 8-22, and 1994-cohort 12-29.

We have two cohorts across four countries, so for simplicity we will build a function so it can be easily replicated.

*LGCA helpers*

```{r lgca-helpers}
#| code-fold: true

# Build lavaan growth syntax from a vector of ages, centred at the first age.
# type: "intercept", "linear", or "conditional"
make_spec <- function(ages, shape = "linear", conditional = FALSE, extra = NULL, covs = c("minority", "chsex"), outcome = "cladder") {
  t    <- (ages - ages[1]) / 10          # decades 
  cols <- paste0(outcome, "_age", ages)
  i    <- paste0("i =~ ", paste0("1*", cols, collapse = " + "))

  if (shape == "intercept") {
    spec <- i; factors <- "i"

  } else if (shape == "basis") {                       # latent-basis (freed-loading)
    load <- rep("NA", length(ages))                    # middle loadings free
    load[1] <- "0"; load[length(ages)] <- "1"          # first fixed 0, last fixed 1
    s <- paste0("s =~ ", paste0(load, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")

  } else {
    s <- paste0("s =~ ", paste0(t, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")
    if (shape %in% c("quadratic", "cubic")) {
      q <- paste0("q =~ ", paste0(t^2, "*", cols, collapse = " + "))
      spec <- paste(spec, q, sep = "\n"); factors <- c(factors, "q")
    }
    if (shape == "cubic") {
      cub <- paste0("cub =~ ", paste0(t^3, "*", cols, collapse = " + "))
      spec <- paste(spec, cub, sep = "\n"); factors <- c(factors, "cub")
    }
  }

  if (conditional) {
    rhs <- paste(covs, collapse = " + ")
    spec <- paste(c(spec, paste0(factors, " ~ ", rhs)), collapse = "\n")
  }
  if(!is.null(extra)){
    spec <- paste(c(spec, extra), collapse ="\n")
  }
  spec
}

# Ages available per cohort per country
ages_comb <- list(ethiopia = c(8,12,15,19,22,26,29), india = c(8,12,15,19,22,26,29),
                peru = c(8,12,15,19,22,26,29), vietnam = c(8,12,15,19,22,26))
ages_yc <- list(ethiopia = c(8,12,15,19,22), india = c(8,12,15,19,22),
                peru = c(8,12,15,19,22), vietnam = c(8,12,15,19)) # vietnam up to age 19 only
ages_oc <- list(ethiopia = c(12,15,19,22,26,29), india = c(12,15,19,22,26,29),
                peru = c(12,15,19,22,26,29), vietnam = c(12,15,19,22,26)) # vietnam up to age 26 only

# Fit one model type for one cohort, across all countries
fit_all <- function(ages_list, shape, cohort_val, exp, conditional = FALSE, constraints = NULL, sex_val = NULL, outcome = "cladder") {
  imap(wide, function(df, ctry) {
    covs <- covs_for(exp)
    if (!is.null(sex_val)) covs <- setdiff(covs, "chsex") # in stratification, drop chsex from conditioning
    spec <- make_spec(ages_list[[ctry]], shape, conditional = conditional,
                      extra = constraints[[ctry]], covs = covs, outcome = outcome) # can change from cladder to careladder easily as needed
    if (!identical(cohort_val, "all")) df <- filter(df, yc == cohort_val)
    if (!is.null(sex_val)) df <- filter(df, chsex == sex_val) # stratify
    keep <- unique(c(exp, covs))
    d <- df |>
      filter(if_all(all_of(keep), ~ !is.na(.x))) |>
      mutate(across(all_of(keep)))
    tryCatch(growth(spec, data = d, missing = "fiml"), error = function(e) NULL)
  })
}

get_fit <- function(fits) {
  extract <- function(f) {
    if (is.null(f)) return(NULL)
    if (!lavInspect(f, "converged")) return(NULL)
    as_tibble(as.list(fitMeasures(f, c(
      "chisq", "df", "pvalue",           # model chi-square
      "cfi", "tli",
      "rmsea", "rmsea.ci.lower", "rmsea.ci.upper",   # RMSEA + its 90% CI
      "srmr",                           
      "aic", "bic", "bic2"               # bic2 = sample-size-adjusted BIC
    ))))
  }
  map_dfr(fits, extract, .id = "country")
}

get_effects <- function(fits) {
  extract <- function(f) {
    if (is.null(f)) return(NULL)
    if (!lavInspect(f, "converged")) return(NULL)
    parameterEstimates(f) |>
      filter(op == "~", lhs %in% c("i", "s", "q", "cub")) |>   
      select(lhs, rhs, est, se, pvalue)
  }
  map_dfr(fits, extract, .id = "country")
}

check_model <- function(fit) {
  if (is.null(fit)) return("NULL (errored)")
  if (!lavInspect(fit, "converged")) return("did not converge")
  pe   <- parameterEstimates(fit)
  vars <- pe[pe$op == "~~" & pe$lhs == pe$rhs, ]     # all variances
  if (any(vars$est < 0)) return("Heywood (negative variance)")
  "OK"
}
```

Code to check Heywood cases: parameterEstimates(model\$country) \|\> dplyr::filter(op=="", lhs==rhs, est \< 0)

------------------------------------------------------------------------

### 1994-cohort age 12-29

#### Check fit for models

Intercept, linear, quadratic, cubic, and basis

[**Religious minority**]{.underline}

```{r oc-religious}

oc_int   <- fit_all(ages_oc, "intercept", 0, exp = "religious_minority")
oc_lin   <- fit_all(ages_oc, "linear",    0, exp = "religious_minority")
oc_quad  <- fit_all(ages_oc, "quadratic", 0, exp = "religious_minority")
oc_cubic <- fit_all(ages_oc, "cubic",     0, exp = "religious_minority")
oc_basis <- fit_all(ages_oc, "basis",     0, exp = "religious_minority")

map(list(int = oc_int, lin = oc_lin, quad = oc_quad, cubic = oc_cubic, basis = oc_basis), get_fit)
```

We use basis model, also known as **freed-loading** model, because trajectories were **not monotonic**. Linear and quadratic models fit poorly, but latent-basis (freed-loading) growth model can estimate the shape of average trajectory from the data rather than imposing a functional form.

In basis models, the first and last factor loadings are fixed to 0 and 1 respectively and in between loadings are freely estimated. Intercept = expected outcome at initial stage; slope = total change across observed age range; estimated loadings = proportion of total change realised at each save.

*Vietnam's data goes up to age 26 only (not age 29), so the basis model is estimated over 5 rather than 6 timepoints, with fewer degrees of freedom. This is why CFI and TLI are 1.00 - boundary values. Fit indices are thus less informative for Vietnam.*

[**Ethnic minority**]{.underline}

```{r oc-ethnic}
oc_int   <- fit_all(ages_oc, "intercept", 0, exp = "ethnic_minority")
oc_lin   <- fit_all(ages_oc, "linear",    0, exp = "ethnic_minority")
oc_quad  <- fit_all(ages_oc, "quadratic", 0, exp = "ethnic_minority")
oc_cubic <- fit_all(ages_oc, "cubic",     0, exp = "ethnic_minority")
oc_basis <- fit_all(ages_oc, "basis",     0, exp = "ethnic_minority")

map(list(int = oc_int, lin = oc_lin, quad = oc_quad, cubic = oc_cubic, basis = oc_basis), get_fit)
```

Basis model fits the best.

[**Sex**]{.underline}

```{r oc-sex}
oc_int   <- fit_all(ages_oc, "intercept", 0, exp = "chsex")
oc_lin   <- fit_all(ages_oc, "linear",    0, exp = "chsex")
oc_quad  <- fit_all(ages_oc, "quadratic", 0, exp = "chsex")
oc_cubic <- fit_all(ages_oc, "cubic",     0, exp = "chsex")
oc_basis <- fit_all(ages_oc, "basis",     0, exp = "chsex")

map(list(int = oc_int, lin = oc_lin, quad = oc_quad, cubic = oc_cubic, basis = oc_basis), get_fit)
```

Basis model still fits the best.

#### Conditional models

For LGCA models for religious minority and ethnic minority, sex is included as a covariate. When the LGCA model is for sex, there are no covariates.

```{r oc-conditional}
#| code-fold: true

oc_basis_cond <- set_names(exposures) |>
  map(~ fit_all(ages_oc, "basis", 0, exp = .x, conditional = TRUE))

# effects with exposure kept as a column
oc_effects <- imap_dfr(oc_basis_cond, ~ get_effects(.x) |> mutate(exposure = .y, .before = 1))

oc_effects |> filter(rhs == exposure, lhs %in% c("i", "s"))   # show only effect of interest per country (not sex as covariate)
```

**Religious minority**: no difference at baseline or in rate of change between religious minorities and majorities over time

**Ethnic/caste minority**

- [India]{.underline}: minorities (compared to Other Castes i.e. higher caste) have lower life satisfaction at baseline (age 12) (-0.94, p\<0.001) but improve more quickly over time (0.47, p=0.02)

- [Vietnam]{.underline}: minorities have lower life satisfaction at baseline (-1.19, p\<0.001) but improve more quickly over time (0.90, p\<0.001)

Importantly this is freed-loading model so overall even though improvement is faster this is the overall change from age 12 to 29 rather than a per-year linear increase.

**Sex**: in [Peru]{.underline}, girls in Peru had higher life satisfaction (intercept b = 0.42, p\<.001) than boys but no difference in change over time.

------------------------------------------------------------------------

### 2001-cohort age 8-22 (replication sample)

[**Religious minority**]{.underline}

```{r yc-religious}
yc_int   <- fit_all(ages_yc, "intercept", 1, exp = "religious_minority")
yc_lin   <- fit_all(ages_yc, "linear",    1, exp = "religious_minority")
yc_quad  <- fit_all(ages_yc, "quadratic", 1, exp = "religious_minority")
yc_cubic <- fit_all(ages_yc, "cubic",     1, exp = "religious_minority")
yc_basis <- fit_all(ages_yc, "basis",     1, exp = "religious_minority", constraints = list(ethiopia = "s ~~ 0*s"))

map(list(int = yc_int, lin = yc_lin, quad = yc_quad, cubic = yc_cubic, basis = yc_basis), get_fit)
```

In Ethiopia 2001-cohort, the slope-factor variance was non-significant and marginally negative (Heywood case), suggesting negligible variability in trajectory of change between individuals. To fit a model we thus have to constrain the slope-factor variance to 0. This only affects estimation of individual differences in slope, not the mean trajectory or estimated minority-majority effects.

Code to check whether models have negative variance still: map_chr(yc_basis, check_model)

[**Ethnic minority**]{.underline}

```{r yc-ethnic}
yc_int   <- fit_all(ages_yc, "intercept", 1, exp = "ethnic_minority")
yc_lin   <- fit_all(ages_yc, "linear",    1, exp = "ethnic_minority")
yc_quad  <- fit_all(ages_yc, "quadratic", 1, exp = "ethnic_minority")
yc_cubic <- fit_all(ages_yc, "cubic",     1, exp = "ethnic_minority")
yc_basis <- fit_all(ages_yc, "basis",     1, exp = "ethnic_minority", constraints = list(ethiopia = "s ~~ 0*s"))

map(list(int = yc_int, lin = yc_lin, quad = yc_quad, cubic = yc_cubic, basis = yc_basis), get_fit)
```

[**Sex**]{.underline}

```{r yc-sex}
yc_int   <- fit_all(ages_yc, "intercept", 1, exp = "chsex")
yc_lin   <- fit_all(ages_yc, "linear",    1, exp = "chsex")
yc_quad  <- fit_all(ages_yc, "quadratic", 1, exp = "chsex")
yc_cubic <- fit_all(ages_yc, "cubic",     1, exp = "chsex")
yc_basis <- fit_all(ages_yc, "basis",     1, exp = "chsex", constraints = list(ethiopia = "s ~~ 0*s"))

map(list(int = yc_int, lin = yc_lin, quad = yc_quad, cubic = yc_cubic, basis = yc_basis), get_fit)
```

#### Conditional models

```{r yc-conditional}
#| code-fold: true

yc_basis_cond <- set_names(exposures) |>
  map(~ fit_all(ages_yc, "basis", 1, exp = .x, conditional = TRUE))

# effects with exposure kept as a column
yc_effects <- imap_dfr(yc_basis_cond, ~ get_effects(.x) |> mutate(exposure = .y, .before = 1))

yc_effects |> filter(rhs == exposure, lhs %in% c("i", "s"))   # show only effect of interest per country (not sex as covariate)
```

*Results for the 2001-cohort replication sample - slightly different from main cohort:*

**Religious minority**

- [Ethiopia]{.underline}: religious minorities similar life satisfaction at baseline (age 12) but improve more quickly over time (0.08, p=0.04)

<!-- -->

- For all other countries, there was no difference at baseline or in rate of change between religious minorities and majorities over time

**Ethnic/caste minority**

- [Ethiopia]{.underline}: similar life satisfaction at baseline (age 12), but minorities improve more quickly over time (0.18, p\<0.001)

- [India]{.underline}: minorities (compared to Other Castes i.e. higher caste) have lower life satisfaction at baseline (age 12) (-0.54, p\<0.001) but similar rate of change over time

- [Vietnam]{.underline}: minorities have lower life satisfaction at baseline (-0.72, p\<0.001) but similar rate of change over time

Importantly this is freed-loading model so overall even though improvement is faster this is the overall change from age 12 to 29 rather than a per-year linear increase.

**Sex**

- [India]{.underline}: girls higher life satisfaction at baseline (-0.21, p\<0.01) than boys but no difference in change over tim

------------------------------------------------------------------------

------------------------------------------------------------------------

### Trajectory visualisations

Visualise model-implied trajectories of minority vs majority groups per country, using basis model.

```{r traj-basis}
#| code-fold: true

implied_traj_basis <- function(fit, ages, exp) {
  if(is.null(fit)) return(NULL)
  if(!lavInspect(fit, "converged")) return(NULL)
  
  est <- parameterEstimates(fit)
  cols <- paste0("cladder_age", ages)
  
  # estimated basis loadings on s (freed loadings), pulled in wave order
  load <- sapply(cols, function(cc)
    est$est[est$op == "=~" & est$lhs == "s" & est$rhs == cc])
  
  cf <- coef(fit)
  ref_i <- cf["i~1"]; ref_s <- cf["s~1"]
  lvl1_i <- ref_i + cf[paste0("i~", exp)]
  lvl1_s <- ref_s + cf[paste0("s~", exp)]
  
  lvl1 <- exp_levels[[exp]][1]; lvl2 <- exp_levels[[exp]][2]
  tibble(
    age = rep(ages, 2),
    group = factor(rep(c(lvl2, lvl1), each = length(ages)), levels = exp_levels[[exp]]),
    fit = c(ref_i + load * ref_s, # majority: intercept + loading x slope
            lvl1_i + load * lvl1_s) # minority: same, shifted by minority coefs
  )
}
```

```{r plot-basis}
#| code-fold: true

plot_obs_model <- function(fits, ages_list, cohort_val, exp, title) {
  model_df <- imap_dfr(fits, function(f, ctry) {
    d <- implied_traj_basis(f, ages_list[[ctry]], exp)
    if (is.null(d)) return(NULL)
    mutate(d, country = ctry)
  })

  obs_df <- imap_dfr(stacked, function(df, ctry) {
    df |> filter(yc == cohort_val, !is.na(cladder), !is.na(.data[[exp]]),
                 equivalent_age %in% ages_list[[ctry]]) |>
      mutate(group = factor(.data[[exp]], levels = exp_levels[[exp]])) |> 
      group_by(country = ctry, age = equivalent_age, group) |>
      summarise(mean_cl = mean(cladder),
                se = sd(cladder) / sqrt(n()),          # <- SE of the mean
                .groups = "drop") |>
      mutate(lo = mean_cl - 1.96 * se,
             hi = mean_cl + 1.96 * se)
  })

  ggplot() +
    geom_ribbon(data = obs_df,
                aes(age, ymin = lo, ymax = hi, fill = group),
                alpha = 0.15, colour = NA) +               # <- the bands
    geom_point(data = obs_df, aes(age, mean_cl, colour = group), alpha = 0.6) +
    geom_line(data = model_df, aes(age, fit, colour = group), linewidth = 1) +
    facet_wrap(~ country) +
    scale_colour_manual(values = exp_cols[[exp]]) +
    scale_fill_manual(values   = exp_cols[[exp]]) +
    labs(title = title, y = "Life satisfaction", x = "Age", colour = NULL, fill = NULL) +
    theme_minimal()
}

walk(exposures, ~ print(
  plot_obs_model(oc_basis_cond[[.x]], ages_oc, 0, exp = .x,
                 title = paste0("1994-Cohort — ", .x))
))
```

Plot 2001-cohort replication sample

```{r plot-basis-yc}
#| code-fold: true

walk(exposures, ~ print(
  plot_obs_model(yc_basis_cond[[.x]], ages_yc, 1, exp = .x,
                 title = paste0("2001-Cohort — ", .x))
))
```

------------------------------------------------------------------------

## Psychological distress models

With SDQ-E we will do both a measurement and structural model together, because the response options are different across ages within the same country, we will construct a latent variable emotional symptoms using the five SDQ items for each timepoint, and moel that.

WLSMV (Weighted Least Squares Mean and Variance) - for CFA here; SDQ-E is ordinal.

### *Measurement model*

*Create latent variable SDQ-E using 5 individual items*

Per wave per country

```{r measurement-model-sdq}
#| code-fold: true

# Confirmatory factor analysis
fit_cfa_wave <- function(df, r) {
  items <- paste0(sdq_items, "_", r)
  spec  <- paste0("emo =~ ", paste(items, collapse = " + "))
  d <- df |> filter(yc == 0)                    # older cohort only
  tryCatch(
    lavaan::cfa(spec, data = d, ordered = items,
                estimator = "WLSMV", missing = "pairwise"),
    error = function(e) NULL)
}

# All countries for all 3 waves
cfa_singlewave <- imap(wide, function(df, ctry) {
  list(age15 = fit_cfa_wave(df, "r3"),
       age19 = fit_cfa_wave(df, "r4"),
       age22 = fit_cfa_wave(df, "r5"))
})
purrr::map(cfa_singlewave, ~ purrr::map_chr(.x, check_model))

# Check fit and loadings
purrr::imap(cfa_singlewave, function(waves, ctry) {
  purrr::map_dfr(waves, function(f) as_tibble(as.list(fitMeasures(f, c("cfi","tli","rmsea","srmr")))),
                 .id = "wave") |> mutate(country = ctry)
})

# Example: Ethiopia age 15 — should all be positive, ideally >.5
standardizedSolution(cfa_singlewave$ethiopia$age15) |>
  dplyr::filter(op == "=~") |> dplyr::select(rhs, est.std, pvalue)
```

*Test longitudinal invariance*

```{r sdq-invariance}
#| code-fold: true

# 3-wave longitudinal CFA at a given invariance level 
# level: "configural" (nothing constrained), "metric" (loadings equal), "scalar "loadings + thresholds equal)
# n_thresh = number of thresholds a country's items have at waves 19/22 - Peru is 3pt i.e. 2 thresholds, remaining are 4pt i.e. 3 thresholds

make_long_cfa <- function(level = "configural", n_thresh = 3) {
  waves <- c("15" = "r3", "19" = "r4", "22" = "r5")
  facs <- paste0("e", names(waves)) # e15, e19, e22
  lines <- character(0)
  
  # measurement model per wave
  for (k in seq_along(waves)) {
    it <- paste0(sdq_items, "_", waves[k])
    labs <- if(level == "configural") paste0("L", seq_along(sdq_items), "_w", k)
            else                      paste0("L", seq_along(sdq_items)) 
    lines <- c(lines, paste0(facs[k], " =~ ", paste0(labs, "*", it, collapse = " + ")))
  }
  
  # full scalar
  if(level == "scalar") {
    for (itm in sdq_items) {
      # constrain shared thresholds across three waves
      i15 <- paste0(itm, "_r3"); i19 <- paste0(itm, "_r4"); i22 <- paste0(itm, "_r5")
      lines <- c(lines,
                 paste0(i15, " | th1*t1"), paste0(i19, " | th1*t1"), paste0(i22, " | th1*t1"),
                 paste0(i15, " | th2*t2"), paste0(i19, " | th2*t2"), paste0(i22, " | th2*t2"))
    }
  }
  if(level == "partial") {
    # partial scalar, free age 15
    for (itm in sdq_items) {
      i19 <- paste0(itm,"_r4"); i22 <- paste0(itm,"_r5")
      for (th in seq_len(n_thresh)) { # 2 for Peru, 3 for others
        lbl <- paste0("th", itm, th)
        lines <- c(lines,
                 paste0(i19, " | ", lbl, "*t", th), 
                 paste0(i22, " | ", lbl, "*t", th)) 
                 # age 15 not listed so can be freely estimated
    }
    }
  }    
  
  # correlated residuals
  for (itm in sdq_items) {
    cols <- paste0(itm, "_", waves)
    lines <- c(lines,
               paste0(cols[1], " ~~ ", cols[2]),
               paste0(cols[1], " ~~ ", cols[3]),
               paste0(cols[2], " ~~ ", cols[3]))
  }
  paste(lines, collapse = "\n")
}

# Fit all four levels for one country; n_thresh set per country
fit_invariance <- function(df, n_thresh = 3) {
  items <- as.vector(outer(sdq_items, c("r3", "r4", "r5"), \(a,b) paste0(a,"_",b)))
  d <- df |> filter(yc==0)
  fit_one <- function(level)
    tryCatch(
    lavaan::cfa(make_long_cfa(level, n_thresh = n_thresh), data = d, ordered = items,
                estimator = "WLSMV", missing = "pairwise",
                parameterization = "theta"),
    error = function(e) NULL)
  list(
    configural = fit_one("configural"),
    metric = fit_one("metric"),
    scalar = fit_one("scalar"),
    partial = fit_one("partial"))
}

# per country threshold count
n_thresh_by_country <- c(ethiopia = 3, india = 3, peru = 2, vietnam = 3)
```

The drop from configural -\> metric is \~0.01 so that's okay. Configural and metric invariance hold - SDQ-E measures the same construct across the three waves.

However, from metric -\> scalar the change is 0.20 i.e. fail; this is expected because of the change in threshold for three countries, from a 3pt to 4pt response. Age 15 SDQ is not comparable to 19/22.

The only solution is to adopt partial scalar invariance, meaning we constrain the thresholds across 19-22 (same scale) but freely estimate the item thresholds at age 15. This retains all three waves while accounting for measurement change. However we still need to interpret the age 15 latent factor with appropriate caution.

For Ethiopia, India, and Vietnam, it makes sense to have partial invariance structure because of change in threshold across rounds. Interesting that despite uniform scaling across time in Peru full scalar did not hold; for consistency across countries we free age-15 for all, including Peru. Though for Peru the reasons are different - perhaps a developmental (age-related) threshold shift where interpretation of the scale differs across ages.

*Partial scalar invariance*

```{r partial-scalar-invariance}
#| code-fold: true

inv_eth <- fit_invariance(wide$ethiopia, n_thresh = n_thresh_by_country["ethiopia"])
inv_ind <- fit_invariance(wide$india,    n_thresh = n_thresh_by_country["india"])
inv_per <- fit_invariance(wide$peru,     n_thresh = n_thresh_by_country["peru"])
inv_vnm <- fit_invariance(wide$vietnam,  n_thresh = n_thresh_by_country["vietnam"])

compare_invariance <- function(inv) {
  ok <- purrr::map_lgl(inv, ~ !is.null(.x) && lavInspect(.x, "converged"))
  inv <- inv[ok]
  purrr::map_dfr(inv, function(f)
    as_tibble(as.list(fitMeasures(f, c("cfi","tli","rmsea","srmr")))),
    .id = "level")
}

# admissibility + the full configural -> metric -> scalar -> partial table per country
list(ethiopia = inv_eth, india = inv_ind, peru = inv_per, vietnam = inv_vnm) |>
  purrr::imap_dfr(function(inv, ctry)
    compare_invariance(inv) |> mutate(country = ctry))
```

Partial invariance recovered well for all which is good news!

### Growth model

Second-order growth model built on partial-invariance measurement structure

```{r growth-model}
#| code-fold: true

# Growth: intercept + linear slope on three wave-factors
make_spec_sdq <- function(shape = "linear", conditional = FALSE,
                          extra = NULL, n_thresh = 3, covs = c("minority", "chsex")) {
  waves <- c("15" = "r3", "19" = "r4", "22" = "r5")
  ages <- c(15, 19, 22)
  facs <- paste0("e", names(waves))
  t <- (ages - ages[1]) / 10
  lines <- character(0)
  
  # measurement: metric loadings
  for (k in seq_along(waves)) {
    it <-  paste0(sdq_items, "_", waves[k])
    labs <- paste0("L", seq_along(sdq_items))
    lines <- c(lines, paste0(facs[k], " =~ ", paste0(labs, "*", it, collapse = " + ")))
  }
  
  # partial scalar thresholds
  for (itm in sdq_items) {
    i19 <- paste0(itm, "_r4"); i22 <- paste0(itm, "_r5")
    for (th in seq_len(n_thresh)) {
      lbl <- paste0("th", itm, th)
      lines <- c(lines, paste0(i19, " | ", lbl, "*t", th),
                        paste0(i22, " | ", lbl, "*t", th))
    }
  }
  
  # correlated residuals
  for (itm in sdq_items) {
    cols <- paste0(itm, "_", waves)
    lines <- c(lines, paste0(cols[1], " ~~ ", cols[2]),
               paste0(cols[1], " ~~ ", cols[3]),
               paste0(cols[2], " ~~ ", cols[3]))
  }
  
  # growth factors on three-wave factors
  i <- paste0("i =~ ", paste0("1*", facs, collapse = " + "))
  if (shape == "intercept") { g <- i; gf <- "i" }
  else {  # linear - only 3 waves
    s <- paste0("s =~ ", paste0(t, "*", facs, collapse = " + "))
    g <- paste(i, s, sep = "\n"); gf <- c("i", "s")
  }
  lines <- c(lines, g)
  
  if (conditional) lines <- c(lines, paste0(gf, " ~ ", paste(covs, collapse = " + ")))
  if(!is.null(extra)) lines <- c(lines, extra)
  paste(lines, collapse = "\n")
}

n_thresh_by_country <- c(ethiopia = 3, india = 3, peru = 2, vietnam = 3)

fit_sdq <- function(shape, exp, conditional = FALSE, constraints = NULL, sex_val = NULL) {
  items <- as.vector(outer(sdq_items, c("r3","r4","r5"), \(a,b) paste0(a,"_",b)))
  imap(wide, function(df, ctry) {                    # loops over 4 countries
    covs <- covs_for(exp)
    if (!is.null(sex_val)) covs <- setdiff(covs, "chsex") # drop chsex in stratification
    keep <- unique(c(exp, covs))
    d <- df |> filter(yc == 0, if_all(all_of(keep), ~ !is.na(.x))) 
    if (!is.null(sex_val)) d <- filter(d, chsex == sex_val) # stratify
    d <- d |> mutate(across(all_of(keep), as.numeric))
    spec <- make_spec_sdq(shape, conditional, extra = constraints[[ctry]],
                           n_thresh = n_thresh_by_country[[ctry]], covs = covs)
    tryCatch(
      lavaan::growth(spec, data = d, ordered = items, estimator = "WLSMV", # weighted least squares mean and variance adjusted 
                     missing = "pairwise", parameterization = "theta"),
      error = function(e) NULL)
  })
}
```

### Structural model

[**Religious minority**]{.underline}

```{r sdq-rel}
sdq_int <- fit_sdq("intercept", exp = "religious_minority")
sdq_lin <- fit_sdq("linear", exp = "religious_minority", constraints = list(
  ethiopia = "i ~~ 0*i", # intercept variance going negative, need to constrain
  india = "s ~~ 0*s")) # slope variance constrained

map_chr(sdq_int, check_model)
map_chr(sdq_lin, check_model)
get_fit(sdq_lin)
```

In Ethiopia, intercept variance was non-significant and marginally negative (Heywood case), so we constrained to 0 - indicating negligible estimable individual variability in baseline SDQ-E. Avoided inadmissible solution. In India, slope variance was similarly constrained, reflecting the limited capacity of a 3-wave model to estimate individual differences in growth factors.

Given we only have 3 waves, a higher-order or freed-loading trajectory shape was not estimable. We specify only a linear slope, and the linear model fits well according to indices - CFI 0.96-0.99, RMSEA \<=0.043, SRMR \~0.05.

[**Ethnic minority**]{.underline}

```{r sdq-ethnic}
sdq_int <- fit_sdq("intercept", exp = "ethnic_minority")
sdq_lin <- fit_sdq("linear", exp = "ethnic_minority", constraints = list(
  ethiopia = "i ~~ 0*i",
  india = "s ~~ 0*s")) 

map_chr(sdq_int, check_model)
map_chr(sdq_lin, check_model)
get_fit(sdq_lin)
```

Linear model fits well.

[**Sex**]{.underline}

```{r sdq-sex}
sdq_int <- fit_sdq("intercept", exp = "chsex")
sdq_lin <- fit_sdq("linear", exp = "chsex", constraints = list(
  ethiopia = "i ~~ 0*i", 
  india = "s ~~ 0*s"))

map_chr(sdq_int, check_model)
map_chr(sdq_lin, check_model)
get_fit(sdq_lin)
```

Linear model still fits well.

*Conditional*

```{r conditional-sdq}
sdq_cond <- set_names(exposures) |> 
  map(~ fit_sdq("linear", exp = .x, conditional = TRUE, constraints = list(
  ethiopia = "i ~~ 0*i",
  india    = "s ~~ 0*s")))

map(sdq_cond, ~ map_chr(.x, check_model))  
    
sdq_effects <- imap_dfr(sdq_cond, ~ get_effects(.x) |> mutate(exposure = .y, .before = 1))
sdq_effects |> filter(rhs == exposure, lhs %in% c("i", "s"))
```

**Religious minority**: for all countries, there were no differences in psychological distress (emotional symptoms) at age 15 or rate of change across 15-22.

**Ethnic minority**

- [Ethiopia:]{.underline} ethnic minorities higher psychological distress at 15 (0.16, p\<0.01) but distress reduces more quickly than majorities (-0.43, p\<0.001)

- [India]{.underline}: ethnic minorities higher distress at age 15 (0.28, p\<0.01) but no difference in rate of change over time

- [Vietnam]{.underline}: ethnic minorities similar distress at baseline but distress increases more quickly across 15-22 (0.40, p=0.02)

**Sex**

- [Ethiopia]{.underline}: similar distress at 15 (-0.10, p=.065), but girls' distress increases more quickly over time (-0.55, p\<0.001)

- [India]{.underline}: girls higher distress at baseline (-0.66, p\<0.001) and no difference in rate of change

- [Peru]{.underline}: boys higher distress at baseline (0.46, p\<0.001) and distress increases more quickly over time (0.49, p=0.001)

- [Vietnam]{.underline}: girls higher distress at baseline (-0.26, p\<0.001) and no difference in rate of change

------------------------------------------------------------------------

------------------------------------------------------------------------

### Trajectory visualisations

*Plot latent factor trajectory*

SDQ is not a latent factor and not observed score, which means we are plotting the model-implied latent factor means at each wave.

**By religious minority/majority**

```{r traj-sdq}
implied_traj_sdq <- function(fit, exp, ages = c(15,19,22)) {
  if (is.null(fit) || !lavInspect(fit, "converged")) return(NULL)
  cf <- coef(fit)
  t <- (ages - ages[1]) / 10 # 0, 0.4, 0.7 - matches slope coding
  
  # intercept and slope means for reference
  ref_i <- cf[["i~1"]]; ref_s <- cf[["s~1"]] 
  lvl1_i <- ref_i + cf[[paste0("i~", exp)]]
  lvl1_s <- ref_s + cf[[paste0("s~", exp)]]
  
  lvl1 <- exp_levels[[exp]][1]; lvl2 <- exp_levels[[exp]][2]
  
  tibble(
    age = rep(ages,2),
    group = factor(rep(c(lvl2, lvl1), each = length(ages)), levels = exp_levels[[exp]]),
    fit = c(ref_i + ref_s * t, lvl1_i + lvl1_s * t)
  )
}
```

```{r plot-sdq}

exp_label <- c(religious_minority = "Religious minority vs majority",
               ethnic_minority = "Racial/ethnic/caste minority vs majority",
               chsex = "Female vs male")

plot_sdq <- function(fits, exp) {
  model_df <- imap_dfr(fits, function(f, ctry) {
    d <- implied_traj_sdq(f, exp)
    if (is.null(d)) return(NULL)
    mutate(d, country = ctry)
  })
  annot <- get_effects(fits) |> 
    filter(lhs == "s", rhs == exp) |> 
    mutate(label = paste0(exp_levels[[exp]][1], " slope p = ", sprintf("%.2f", pvalue)))
  
  ggplot(model_df, aes(age, fit, colour = group)) +
    geom_line(linewidth = 1) + 
    geom_point() +
    geom_text(data = annot, aes(x = 15, y = Inf, label = label), 
              colour = "grey30", hjust = 0, vjust = 1.4, size = 3, inherit.aes = FALSE) +
    facet_wrap(~ country) +
    scale_colour_manual(values = exp_cols[[exp]]) +
    scale_x_continuous(breaks = c(15, 19, 22)) +
    labs(title = "Psychological distress (SDQ-E) model-implied latent trajectories",
         subtitle = paste0(exp_levels[[exp]],
                           if (exp != "chsex") " - adjusted for sex" else ""),,
         caption = "Higher values indicate greater distress. Trajectories are model-implied estimates.",
         y = "Latent psychological distress (model-implied)", x = "Age", colour = NULL) +
    theme_minimal()
}

walk(exposures, ~ print(plot_sdq(sdq_cond[[.x]], .x)))
```

The y-axis is latent factor scores so they are not interpretable SDQ scores - you only interpret the differences, not absolute levels.

------------------------------------------------------------------------

------------------------------------------------------------------------

## Sex stratification

### Life satisfaction

1994-cohort only

```{r stratify-sex}
oc_male <- fit_all(ages_oc, "basis", 0, conditional = TRUE, sex_val = 0)
oc_female <- fit_all(ages_oc, "basis", 0, conditional = TRUE, sex_val = 1,
                     constraints = list(
                       india = "s ~~ 0*s",
                       vietnam = "s ~~ 0*s",
                       peru = "s ~~ 0*s"))

purrr::map_chr(oc_male, check_model)
purrr::map_chr(oc_female, check_model)
get_effects(oc_male)    
get_effects(oc_female) 
```

Plot

```{r oc-sex-stratified-plot}
implied_traj_strat <- function(fit, ages, sex_label) {
  if (is.null(fit) || !lavInspect(fit, "converged")) return(NULL)
  est <- parameterEstimates(fit)
  cols <- paste0("cladder_age", ages)
  load <- sapply(cols, function(cc)
    est$est[est$op == "=~" & est$lhs == "s" & est$rhs == cc])
  
  cf <- coef(fit)
  maj_i <- cf[["i~1"]]; maj_s <- cf[["s~1"]]
  min_i <- maj_i + cf[["i~minority"]]
  min_s <- maj_s + cf[["s~minority"]]
  
  tibble(
    age = rep(ages, 2),
    group = rep(c("Majority", "Minority"), each = length(ages)),
    sex = sex_label,
    fit = c(maj_i + load * maj_s, min_i + load * min_s)
  )
}

# male/female x majority/minority
strat_df <- bind_rows(
  imap_dfr(oc_male, \(f, ctry) {
    d <- implied_traj_strat(f, ages_oc[[ctry]], "Male"); if(!is.null(d)) mutate(d, country=ctry) }),
  imap_dfr(oc_female, \(f, ctry) {
    d <- implied_traj_strat(f, ages_oc[[ctry]], "Female"); if(!is.null(d)) mutate(d, country=ctry) })
) |> 
  mutate(line = paste(sex, ",", group)) # e.g. "female, minority"

ggplot(strat_df, aes(age, fit, colour = sex, linetype = group)) +
  geom_line(linewidth = 0.9) + geom_point(size = 1) +
  facet_wrap(~country) + 
  scale_colour_manual(values = c("Male" = "#009E73", "Female" = "#E69F00")) + 
  scale_linetype_manual(values = c("Majority" = "solid", "Minority" = "dashed")) +
  guides(linetype = guide_legend(override.aes = list(linewidth = 0.5)),
         colour = guide_legend(override.aes = list(linetype = "solid"))) +
  labs(title = "Life satisfaction: minority vs majority, by sex",
       subtitle = "Sex-stratified latent-basis trajectories",
       y = "Life satisfaction", x = "Age", colour = NULL, linetype = NULL) +
  theme_minimal()
```

2001-cohort only

```{r stratify-sex-yc}
yc_male <- fit_all(ages_yc, "basis", 1, conditional = TRUE, sex_val = 0, constraints = list(vietnam = "s ~~ 0*s"))
yc_female <- fit_all(ages_yc, "basis", 1, conditional = TRUE, sex_val = 1, constraints = list(ethiopia = "s ~~ 0*s", peru = "s ~~ 0*s"))
# Constraints since variance of slope were negative

purrr::map_chr(yc_male, check_model)
purrr::map_chr(yc_female, check_model)
get_effects(yc_male)    
get_effects(yc_female) 
```

```{r yc-sex-stratified-plot}
# male/female x majority/minority
strat_df <- bind_rows(
  imap_dfr(yc_male, \(f, ctry) {
    d <- implied_traj_strat(f, ages_yc[[ctry]], "Male"); if(!is.null(d)) mutate(d, country=ctry) }),
  imap_dfr(yc_female, \(f, ctry) {
    d <- implied_traj_strat(f, ages_yc[[ctry]], "Female"); if(!is.null(d)) mutate(d, country=ctry) })
) |> 
  mutate(line = paste(sex, ",", group)) # e.g. "female, minority"

ggplot(strat_df, aes(age, fit, colour = sex, linetype = group)) +
  geom_line(linewidth = 0.9) + geom_point(size = 1) +
  facet_wrap(~country) + 
  scale_colour_manual(values = c("Male" = "#009E73", "Female" = "#E69F00")) + 
  scale_linetype_manual(values = c("Majority" = "solid", "Minority" = "dashed")) +
  guides(linetype = guide_legend(override.aes = list(linewidth = 0.5)),
         colour = guide_legend(override.aes = list(linetype = "solid"))) +
  labs(title = "Life satisfaction: minority vs majority, by sex (2001-cohort replication test)",
       subtitle = "Sex-stratified latent-basis trajectories",
       y = "Life satisfaction", x = "Age", colour = NULL, linetype = NULL) +
  theme_minimal()
```

### Psychological distress

1994-cohort only

```{r stratify-sex-sdq}
oc_male_sdq <- fit_sdq("linear", conditional = TRUE, sex_val = 0, constraints = list(india = "s ~~ 0*s"))
oc_female_sdq <- fit_sdq("linear", conditional = TRUE, sex_val = 1, constraints = list(ethiopia = "i ~~ 0*i"))

purrr::map_chr(oc_male_sdq, check_model)
purrr::map_chr(oc_female_sdq, check_model)
get_effects(oc_male_sdq)    
get_effects(oc_female_sdq) 
```

The results from sex stratification are the same – for both boys and girls, there are no differences in life satisfaction (Cantril ladder) nor psychological distress (SDQ-E) between religious minorities and majorities.

------------------------------------------------------------------------

------------------------------------------------------------------------

# Caregiver mental health

### Summary

```{r test-careladder}
# Pivot longer, mean/SD/n per country x age x minority group
desc <- all_df |> 
  pivot_longer(careladder, names_to = "outcome", values_to = "value") |> 
  filter(!is.na(value), !is.na(minority)) |> 
  group_by(country, outcome, equivalent_age, minority) |> 
  summarise(mean = mean(value), sd = sd(value), n = n(), .groups = "drop")

# Clear wide table - maj/min side by side
desc_table <- desc |> 
  mutate(cell = sprintf("%.2f (%.2f) [n=%d]", mean, sd, n)) |> 
  select(country, outcome, equivalent_age, minority, cell) |> 
  pivot_wider(names_from = minority, values_from = cell) |> 
  arrange(country, outcome, equivalent_age)

desc_table
```

### Visualise

```{r careladder-traj-min-maj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_careladder_trajectory_minmaj <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(careladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age, minority) |>
    summarise(mean_cl = mean(careladder),
              se = sd(careladder)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(5, 8, 12, 15)) +
    scale_y_continuous(limits = c(2, 7)) +
    labs(title = ctry, x = "Age", y = "Mean caregiver life satisfaction") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = minority), alpha = 0.2, colour = NA) +
    scale_colour_manual(values = group_cols) +
    scale_fill_manual(values = group_cols) +
    theme_minimal()
}

plots <- imap(stacked, plot_careladder_trajectory_minmaj)
wrap_plots(plots, ncol = 2)
```

By cohort

```{r careladder-traj-cohort}
#| code-fold: true

# Mean careladder by age and cohort, at overlap ages, per country
cohort_check <- map(stacked, ~ .x |>
                      filter(!is.na(careladder), equivalent_age %in% c(5, 8, 12, 15)) |>
                      mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
                      group_by(equivalent_age, cohort) |>
                      summarise(mean_cl = mean(careladder), se = sd(careladder)/sqrt(n()), n = n(),
                                .groups = "drop"))

# Plot 
bind_rows(stacked, .id = "country") |>
  filter(!is.na(careladder), equivalent_age %in% c(5, 8, 12, 15)) |>
  mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
  group_by(country, equivalent_age, cohort) |>
  summarise(mean_cl = mean(careladder), se = sd(careladder)/sqrt(n()), .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = cohort, fill = cohort)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  facet_wrap(~ country) +
  scale_x_continuous(breaks = c(5, 8, 12, 15)) +
  scale_y_continuous(limits = c(1,9)) +
  labs(y = "Mean caregiver life satisfaction", x = "Age") +
  theme_minimal()
```

Data mostly from younger cohort, for LGCA we will have to remove older cohort from analysis because there is no change over time we can examine.

### T Test

Keep both cohorts for simple regression to see overall effect.

```{r t-test-careladder}
# Welch t-tests per age for careladder, per country (mirror cladder)
care_ttests <- imap_dfr(wide, function(df, ctry) {
  d <- df |> filter(yc == 1, !is.na(minority))
  map_dfr(c(5, 8, 12, 15), function(a) {
    col <- paste0("careladder_age", a)
    maj <- d[[col]][d$minority == 0]
    min <- d[[col]][d$minority == 1]
    tt <- t.test(min, maj)   # Welch by default
    tibble(age = a, diff = diff(rev(tt$estimate)),
           ci_low = tt$conf.int[1], ci_high = tt$conf.int[2],
           p = tt$p.value, country = ctry)
  })
})
care_ttests
```

### LGCA

*Filter careladder* to younger cohort only and those with 3 complete careladder) for LGCA

```{r filter-sample-careladder}
#| code-fold: true

wide <- map(wide, ~ .x |> 
              mutate(yc == 1) |> 
              mutate(n_cl = rowSums(!is.na(across(starts_with("careladder_age"))))) |> 
              filter(n_cl >= 3) |> 
              select(-n_cl))
```

*Sample size*

```{r sample-size-careladder}
careladder_n <- imap_dfr(wide, function(df, ctry) {
  d <- df |> filter(yc == 1, !is.na(minority), !is.na(chsex))
  tibble(country = ctry, n = nrow(d))
})
careladder_n
```

*LGCA helper*

```{r helper-careladder}
# make_spec edited to include different outcome flexible

# specify ages
ages_careladder <- list(ethiopia = c(5,8,12,15), india = c(5,8,12,15),
                  peru = c(5,8,12,15), vietnam = c(5,8,12,15))
```

*Linear*

```{r lgca-linear-careladder}
care_lin <- fit_all(ages_careladder, "linear", 1, outcome = "careladder")
purrr::map_chr(care_lin, check_model)
get_fit(care_lin)
```

Acceptable but not great fit. Vietnam good; others CFI 0.86-0.90 and RMSEA 0.08-0.09. Try quadratic next.

*Quadratic*

```{r lgca-quadratic-careladder}
care_quad <- fit_all(ages_careladder, "quadratic", 1, outcome = "careladder", constraints = list(
  india = "q ~~ 0*q", vietnam = "q ~~ 0*q\ns ~~ 0*s"))
purrr::map_chr(care_quad, check_model)
get_fit(care_quad)
```

Quadratic fits substantially better than linear, but only when we constrained q in India and q & s in Vietnam. Finally, let's try basis.

*Basis*

```{r lgca-basis-careladder}
care_basis <- fit_all(ages_careladder, "basis", 1, outcome = "careladder")
purrr::map_chr(care_basis, check_model)
get_fit(care_basis)
```

Basis has the best fit of all! It fits cleanly without the constraints quadratic needed. We will use basis in conditional then.

*Conditional*

```{r lgca-conditional-careladder}
care_cond <- fit_all(ages_careladder, "basis", 1, conditional = TRUE, outcome = "careladder")
purrr::map_chr(care_cond, check_model)
get_effects(care_cond)
```

The story for caregivers is more interesting than for their children!

In Ethiopia, religious minority caregivers had lower life satisfaction at baseline (when child is aged 5) (intercept b=-0.388, p\<.001) though in the next decade they improve more quickly than majorities (slope b=0.788, p\<.001).

In India, caregivers from either group start off similar when child is aged 5 but over the next 10 years religious minority caregivers' life satisfaction improve more slowly than majorities (slope b=-0.298, p=0.011).

*Plot*

```{r traj-basis-careladder}
implied_traj_basis_careladder <- function(fit, ages) {
  if(is.null(fit)) return(NULL)
  if(!lavInspect(fit, "converged")) return(NULL)
  
  est <- parameterEstimates(fit)
  cols <- paste0("careladder_age", ages)
  
  # estimated basis loadings on s (freed loadings), pulled in wave order
  load <- sapply(cols, function(cc)
    est$est[est$op == "=~" & est$lhs == "s" & est$rhs == cc])
  
  cf <- coef(fit)
  maj_i <- cf["i~1"]; maj_s <- cf["s~1"]
  min_i <- maj_i + cf["i~minority"]
  min_s <- maj_s + cf["s~minority"]
  
  tibble(
    age = rep(ages, 2),
    group = rep(c("Majority", "Minority"), each = length(ages)),
    fit = c(maj_i + load * maj_s, # majority: intercept + loading x slope
            min_i + load * min_s) # minority: same, shifted by minority coefs
  )
}
```

```{r plot-basis-careladder}
plot_obs_model_careladder <- function(fits, ages_list, cohort_val, title) {
  model_df <- imap_dfr(fits, function(f, ctry) {
    d <- implied_traj_basis_careladder(f, ages_list[[ctry]])
    if (is.null(d)) return(NULL)
    mutate(d, country = ctry)
  })
  
  obs_df <- imap_dfr(stacked, function(df, ctry) {
    df |> filter(yc == cohort_val, !is.na(careladder),
                 equivalent_age %in% ages_list[[ctry]]) |>
      mutate(group = if_else(minority == "Minority", "Minority", "Majority")) |>
      group_by(country = ctry, age = equivalent_age, group) |>
      summarise(mean_cl = mean(careladder),
                se = sd(careladder) / sqrt(n()),          # <- SE of the mean
                .groups = "drop") |>
      mutate(lo = mean_cl - 1.96 * se,
             hi = mean_cl + 1.96 * se)
  })
  
  ggplot() +
    geom_ribbon(data = obs_df,
                aes(age, ymin = lo, ymax = hi, fill = group),
                alpha = 0.15, colour = NA) +               # <- the bands
    geom_point(data = obs_df, aes(age, mean_cl, colour = group), alpha = 0.6) +
    geom_line(data = model_df, aes(age, fit, colour = group), linewidth = 1) +
    facet_wrap(~ country) +
    scale_colour_manual(values = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
    scale_fill_manual(values   = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
    labs(title = title, y = "Caregiver life satisfaction", x = "Child's age", colour = NULL, fill = NULL) +
    theme_minimal()
}

plot_obs_model_careladder(care_cond, ages_careladder, 1, "2001-Cohort Caregivers")
```

**Ethiopia**: religious minority caregivers had lower life satisfaction at baseline (when child is aged 5) (intercept b=-0.388, p\<.001) though in the next decade they improve more quickly than majorities (slope b=0.788, p\<.001).

**India**: caregivers from either group start off similar when child is aged 5 but over the next 10 years religious minority caregivers' life satisfaction improve more slowly than majorities (slope b=-0.298, p=0.011).
