---
title: "Mental health trajectories of religious majority and minority youth: data pre-processing"
format: 
  html: 
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
    df-print: paged
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

This Quarto document pre-processes data from the Young Lives project, accessed via the UK Data Service on 4th August, 2026.

# Set up

```{r setup}
#| include: false

# Install any packages that aren't there yet then load
pacman::p_load(
  tidyverse, # data wrangling
  here, # clean file paths
  lavaan, # latent growth curve models
  haven, 
  patchwork
)
```

```{r load-data}
# here() builds a path from project root, so it works regardless of machine or subfolder
countries <- c("ethiopia", "india", "peru", "vietnam")

# Load rounds 1-7 constructed files 
data_list <- map(countries, ~ read_sav(here("data", paste0(.x, ".sav")))) |>
  set_names(countries)

# Load SDQ-E files (rounds 3-5, 1994-cohort only)
rounds    <- 3:5
combos <- tidyr::expand_grid(country = countries, round = rounds) # every country–round combination
sdq_list <- map2(
  combos$country, combos$round,
  ~ read_sav(here("data", paste0(.x, "_r", .y, "_sdq.sav")))
) |>
  set_names(paste0(combos$country, "_r", combos$round, "_sdq"))

# Note: Generally, the SDQ files for rounds 3-5 are named occh_older_child. The only exception is for Peru rounds 4 and 5, the files are named olderchild-SAQ. 
```

### Variable transformations

Key variables from each file

- Constructed files (ethiopia, india, peru, vietnam): child's ID (*childid*), cohort (*yc*), round, cluster ID (*clustid),* age in years (*ageyea*)*,* **sex** (*chsex*), child's ethnic group (*chethnic*), **child's religious affiliation** (*chldrel*), **life satisfaction** (*cladder*)

- SDQ-E files (ethiopia_r3-5, india_r3-5, peru_r3-5, vietnam_r3-5): strengths and difficulties questionnaire emotional subscale - **psychological distress** (comprised of 5 items) collected at rounds 3, 4 and 5 for the 1994-cohort at ages 15, 19, and 22 respectively

#### Select variables

```{r main-file-variables}

# Variables from constructed file
keep_vars <- c("childid", "yc", "round", "ageyea", "chldrel", "cladder", "chsex", "chethnic", "clustid")
data <- map(data_list, function(df) {
  df |> 
    select(all_of(keep_vars)) |> 
    mutate(childid = as.character(childid))
})
```

```{r sdq-variables}
## Country prefixes for childid
prefixes <- c(ethiopia = "ET", india = "IN", peru = "PE", vietnam = "VN")

## Fix childid: make sure there are 6 digits and add country prefix
fix_childid <- function(id, prefix) {
  id <- as.character(id)
  ifelse(str_detect(id, "^[A-Za-z]"),                 # if already full format
         id,                                          # leave it alone
         paste0(prefix, str_pad(id, 6, "left", "0"))) # otherwise reconstruct
}

## SDQ items come in two naming schemes across countries/rounds
scheme_headac <- function(r) c(
  headache = paste0("HEADACR", r),
  worry    = paste0("WRYLOTR", r),
  unhappy  = paste0("UNHPPYR", r),
  nervous  = paste0("NRVSITR", r),
  fears    = paste0("MNYFERR", r)
)
scheme_feay <- function(r) c(
  headache = paste0("FEAY13R", r),
  worry    = paste0("FEAY05R", r),
  unhappy  = paste0("FEAY20R", r),
  nervous  = paste0("FEAY38R", r),
  fears    = paste0("FEAY28R", r)
)

## Which scheme each country uses at R4/R5 (R3 is same across countries)
country_scheme <- list(
  india    = scheme_feay,
  peru     = scheme_headac,
  ethiopia = scheme_feay,
  vietnam  = scheme_feay
)

## Clean SDQ files: select + rename items and ID, then fix childid
sdq <- imap(sdq_list, function(df, nm) {
  ctry <- str_extract(nm, "^[a-z]+")
  rnum <- str_extract(nm, "[3-5]")
  pfx  <- prefixes[[ctry]]
  
  df <- df |> rename_with(toupper)
  
  # R3 uses HEADAC everywhere; R4/R5 use the country's scheme
  items <- if (rnum == "3") scheme_headac(rnum) else country_scheme[[ctry]](rnum)
  items <- setNames(items, paste0(names(items), "_r", rnum))
  
  idc <- if (rnum == "3") "CHILDID" else "CHILDCODE"
  
  df |>
    select(all_of(c(childid = idc, items))) |>
    mutate(childid = fix_childid(childid, pfx))
})
```

#### Rescale SDQ

```{r inspect-sdq}
#| include: false
## Distinct response values in each country-round, pre-rescale
imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  sort(unique(as.numeric(unlist(df[cols]))))
})
```

SDQ-E needs to be **rescaled and standardised** as measurement was different across rounds and countries.

- Round 3: 1 - not true for you, 2 - a little true for you, 3 - certainly true for you, 77 - NK, 88 - N/A, 99 - Missing

- Rounds 4 and 5: 1 - strongly disagree, 2 - disagree, 3 - agree, 4 - strongly agree, 77 - NK, 79 - refused to answer

SDQ sub-scale is normally **0 - not true, 1 - somewhat true, 2 - certainly true -\> 10 full score**.

Therefore, scale of 1-3 (round 3) will be rescaled to 0-2. As for scale of 1-4 (rounds 4-5), 1,2 will be collapsed to 0 (not true), 3\>1, and 4-\>2.

```{r rescale-sdq}
## --- SDQ rescaling + imputation across all country-round files ---

# Rescale function; unrecognised values (77/79/88/99) fall all become NA
rescale_3pt <- function(x) case_when(x == 1 ~ 0, x == 2 ~ 1, x == 3 ~ 2, TRUE ~ NA_real_)
rescale_4pt <- function(x) case_when(x %in% c(1, 2) ~ 0, x == 3 ~ 1, x == 4 ~ 2, TRUE ~ NA_real_)

# Format at R4/R5 (R3 is 3pt for everyone) — all four confirmed against the data
format_r45 <- c(india = "4pt", peru = "3pt", ethiopia = "4pt", vietnam = "4pt")

sdq <- imap(sdq, function(df, nm) {
  ctry <- str_extract(nm, "^[a-z]+")
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  
  fmt     <- if (rnum == "3") "3pt" else format_r45[[ctry]]
  rescale <- if (fmt == "3pt") rescale_3pt else rescale_4pt
  
  df |>
    mutate(across(all_of(cols), ~ rescale(as.numeric(.x)))) |>
    mutate(
      nonmiss = rowSums(!is.na(across(all_of(cols)))),
      !!paste0("emo_r", rnum) := if_else(
        nonmiss > 2, # need to have answered 3 or more out of 5 items to be included
        round(rowMeans(across(all_of(cols)), na.rm = TRUE) * 5),
        NA_real_
      )
    ) |>
    select(-nonmiss)
})
```

```{r check-sdq}
## Check each round's total emotional score is within 0-10 with sensible NA counts
imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  summary(df[[paste0("emo_r", rnum)]])
})
```

#### Transform variables

```{r check-variables}
# Check distinct values of each key variable 
map(data, function(df) {
  map(df[c("cladder", "chsex", "chldrel", "chethnic", "ageyea")], ~sort(unique(as.numeric(.x))))
})
# No NA values 77/88/99 to transform

# Transform into named categories
map(data, ~ haven::as_factor(.x$chldrel) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chsex) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chethnic) |> table(useNA = "always"))
```

Create separate dataframes for each country

```{r country-dataframes}

# SDQ: one row per child 
sdq_scores <- imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  df |> select(childid, !!paste0("emo_r", rnum))
})

# Merge constructed data and SDQ for each country 
countries <- c("ethiopia", "india", "peru", "vietnam")

merged_data <- map(countries, function(ctry) {
  base      <- data[[ctry]]
  sdq_files <- sdq_scores[str_starts(names(sdq_scores), paste0(ctry, "_"))]
  reduce(sdq_files, left_join, by = "childid", .init = base)
}) |>
  set_names(countries)

# Row counts shouldn't jump (a jump = duplicate childids somewhere)
map_int(merged_data, nrow)

# Create separate dataframes for each country
list2env(merged_data, envir = .GlobalEnv) 
```

### Structuring data

#### Create variable 'equivalent age'

Although we have age in years as the variable for real age, we need an equivalent age of each round i.e. in round 3 for example, the 2001-cohort is aged 8 and the 1994-cohort is aged 15.

```{r equivalent-age}
age_yc <- c("1"=1, "2"=5, "3"=8, "4"=12, "5"=15, "6"=19, "7"=22)
age_oc <- c("1"=8, "2"=12, "3"=15, "4"=19, "5"=22, "6"=26, "7"=29)

add_age <- function(df) {
  df %>%
    mutate(equivalent_age = if_else(yc == 1, age_yc[as.character(round)],
                                    age_oc[as.character(round)])) %>%
    filter(!is.na(equivalent_age)) %>%
    group_by(childid, equivalent_age) %>%
    slice_max(!is.na(cladder), n = 1, with_ties = FALSE) %>% # In round 6 there are multiple rows for phone calls but we only keep the row that has cladder data
    ungroup()
}

stacked <- map(merged_data, add_age)
```

```{r stack-sdq}
# There are three variables for SDQ at the moment, emo_r3, emo_r4 and emo_r5 but the data is long so there should be only one variable and the data split across rounds.

stacked <- map(stacked, \(df) {
  df |> 
    mutate(sdq = case_when(
      round == 3 ~ emo_r3,
      round == 4 ~ emo_r4,
      round == 5 ~ emo_r5,
      TRUE ~ NA_real_
    )) |> 
    select(-emo_r3, -emo_r4, -emo_r5)
})
```

# Categorise minority and majority status

::: callout-note
**Update**: It appears it would be more appropriate to take into account country context when classifying majority/minority rather than solely rely on World Religion Database.
:::

We use religious demographics from 2000 accessed via the World Religion Database for each of the four countries. We only include religions \>1% of the population share.

- **Ethiopia**: 58% Christian, 34% Muslim, 7.0% Ethnic religionists

- **India**: 75% Hindus, 13% Muslims, 4.2% Christians, 3.3% Ethnic religionists, 1.8% Sikhs, 1.4% Agnostics

- **Peru**: 96% Christian, 1.4% Ethnic religionist, 1.2% Agnostic

- **Vietnam**: 48% Buddhist, 13% Agnostic, 12% Ethnic religionist, 10.8% New religionist, 8.3% Christian, 6.7% Atheist, 1.0% Chinese folk religionist

```{r check-religions}
map(stacked, ~ count(.x, chldrel = as_factor(chldrel)))
```

It appears that in Vietnam, although WRD had 48% Buddhist and around 20% non-religious, in the Young Lives sample 90% have no religion. I will put both Buddhist and none as 'majority religion' for now until we make a decision.

```{r classify}

# Specify majority religions
majority_religion <- list(ethiopia = "orthodox", india = "hindu", peru = "catholic", vietnam = c("buddhist", "none", "ancestor worship"))

classify <- function(df, ctry) {
  maj <- majority_religion[[tolower(ctry)]]
  df |> 
    mutate(
      chldrel = as.character(as_factor(chldrel)),
     minority = case_when(
       is.na(chldrel) ~ NA_character_,
       tolower(chldrel) %in% maj ~ "Majority",
       TRUE ~ "Minority"
     )
      )
}

stacked <- imap(stacked, classify)
```

*Later we will create another variable majority_religion_region such that instead of country-level demographics we will use regional demographics to determine majority and minority status*

```{r check-classifications}
map(stacked, ~ .x |>
      distinct(childid, chldrel, minority) |>
      count(chldrel, minority))

map(stacked, ~ .x |>
      distinct(childid, chldrel, minority) |>
      count(minority) |>
      mutate(pct = round(100 * n / sum(n), 1)))

# Check how many NAs, then drop
stacked <- imap(stacked, classify) |> map(~ filter(.x, !is.na(minority)))
```

Missingness for religion (*chldrel*)

- Ethiopia: 2

- India: 0

- Peru: 1

- Vietnam: 1

# Descriptives

## Visualisations

### Life satisfaction - combined cohorts

Overall trajectory

```{r overall-traj}
# Mean cladder by age, overall and by group, per country
plot_trajectory <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(4, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory)
wrap_plots(plots, ncol = 2)
```

By minority/majority status

```{r traj-min-maj}


# Mean cladder by age, overall and by group, per country
plot_trajectory_minmaj <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age, minority) |>
    summarise(mean_cl = mean(cladder),
              se = sd(cladder)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(4, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = minority), alpha = 0.2, colour = NA) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_minmaj)
wrap_plots(plots, ncol = 2)
```

By sex

```{r traj-sex}

# Mean cladder by age, overall and by group, per country
plot_trajectory_sex <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    mutate(chsex = as_factor(chsex)) |> 
    group_by(equivalent_age, chsex) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = chsex)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(3, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sex)
wrap_plots(plots, ncol = 2)
```

By majority/minority status and sex

```{r traj-sex-minority}
bind_rows(stacked, .id = "country") |>
  filter(!is.na(cladder), !is.na(equivalent_age)) |>
  mutate(chsex = as_factor(chsex)) |>
  group_by(country, equivalent_age, chsex, minority) |>
  summarise(mean_cl = mean(cladder),
            se = sd(cladder) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = minority, fill = minority)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96 * se, ymax = mean_cl + 1.96 * se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  facet_grid(country ~ chsex) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  scale_y_continuous(limits = c(4, 7)) +
  theme_minimal()
```

### Psychological distress (SDQ-E) - 1994-cohort only

Overall trajectory

```{r overall-traj-sdq}
# Mean SDQ-E by age, overall and by group, per country
plot_trajectory_sdq <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(sdq), !is.na(equivalent_age)) |>
    group_by(equivalent_age) |>
    summarise(mean_cl = mean(sdq), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(title = ctry, x = "Age", y = "Mean psychological distress") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sdq)
wrap_plots(plots, ncol = 2)
```

By minority/majority status

```{r traj-min-maj-sdq}

# Mean SDQ by age, overall and by group, per country
plot_trajectory_minmaj_sdq <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(sdq), !is.na(equivalent_age)) |>
    group_by(equivalent_age, minority) |>
    summarise(mean_cl = mean(sdq),
              se = sd(sdq)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(title = ctry, x = "Age", y = "Mean psychological distress") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = minority), alpha = 0.2, colour = NA) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_minmaj_sdq)
wrap_plots(plots, ncol = 2)
```

By sex

```{r traj-sex-sdq}

# Mean SDQ by age, overall and by group, per country
plot_trajectory_sex_sdq <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(sdq), !is.na(equivalent_age)) |>
    mutate(chsex = as_factor(chsex)) |> 
    group_by(equivalent_age, chsex) |>
    summarise(mean_cl = mean(sdq), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = chsex)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(title = ctry, x = "Age", y = "Mean psychological distress") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sex_sdq)
wrap_plots(plots, ncol = 2)
```
