---
title: "preprocessing"
format: 
  html: 
    toc: true
    toc-depth: 3
    number-sections: true
    code-fold: false
    df-print: paged
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

# Pre-processing

This Quarto document pre-processes data from the Young Lives project, accessed via the UK Data Service.

### Set-up

```{r setup}
#| label: setup
#| include: false

# Install any packages that aren't there yet then load
pacman::p_load(
  tidyverse, # data wrangling
  here, # clean file paths
  lavaan, # latent growth curve models
  haven
)

#| label: load-data
# here() builds a path from project root, so it works regardless of machine or subfolder

# Load rounds 1-7 constructed files 
ethiopia <- read_sav(here("data", "ethiopia.sav"))
india    <- read_sav(here("data", "india.sav"))
peru     <- read_sav(here("data", "peru.sav"))
vietnam  <- read_sav(here("data", "vietnam.sav"))

# Load SDQ-E files (rounds 3-5, 1994-cohort only)
```

Check data structure

```{r view-files}
# ethiopia |> view()
# india |> view()
# peru |> view()
# vietnam |> view()
```

Analytical variables in each file

- Constructed files (ethiopia, india, peru, vietnam): **child's religious affiliation** (*chldrel*), **life satisfaction** (*cladder*), **sex** (*chsex*), **cluster ID** (*clustid)*

- SDQ-E files (ethiopia_r3-5, india_r3-5, peru_r3-5, vietnam_r3-5): strengths and difficulties questionnaire emotional subscale - **psychological distress** (comprised of 5 items) collected at rounds 3, 4 and 5 for the 1994-cohort at ages 15, 19, and 22 respectively

### Categorise minority and majority status

We use religious demographics from 2000 accessesd via the World Religion Database for each of the four countries. We only include religions \>1% of the population share.

- **Ethiopia**: 58% Christian, 34% Muslim, 7.0% Ethnic religionists

- **India**: 75% Hindus, 13% Muslims, 4.2% Christians, 3.3% Ethnic religionists, 1.8% Sikhs, 1.4% Agnostics

- **Peru**: 96% Christian, 1.4% Ethnic religionist, 1.2% Agnostic

- **Vietnam**: 48% Buddhist, 13% Agnostic, 12% Ethnic religionist, 10.8% New religionist, 8.3% Christian, 6.7% Atheist, 1.0% Chinese folk religionist
