---
title: "Mental health trajectories of religious majority and minority youth: data pre-processing"
format: 
  html: 
    toc: true
    toc-depth: 3
    number-sections: true
    df-print: paged
    embed-resources: true
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

This Quarto document pre-processes data from the Young Lives project, accessed via the UK Data Service on 4th August, 2026.

# Set up

*Load packages*

```{r setup}
#| code-fold: true

# Install any packages that aren't there yet then load
pacman::p_load(
  tidyverse, # data wrangling
  here, # clean file paths
  lavaan, # latent growth curve models
  haven, 
  patchwork, 
  purrr,
  ggplot2
)

# Set colours
group_cols <- c(Majority = "#4A90D9",   # blue
                Minority = "#E05B5B")   # red

options(scipen = 999)
```

*Load data*

```{r load-data}
#| code-fold: true

# here() builds a path from project root, so it works regardless of machine or subfolder
countries <- c("ethiopia", "india", "peru", "vietnam")

# Load rounds 1-7 constructed files 
data_list <- map(countries, ~ read_sav(here("data", paste0(.x, ".sav")))) |>
  set_names(countries)

# Load SDQ-E files (rounds 3-5, 1994-cohort only)
rounds    <- 3:5
combos <- tidyr::expand_grid(country = countries, round = rounds) # every country–round combination
sdq_list <- map2(
  combos$country, combos$round,
  ~ read_sav(here("data", paste0(.x, "_r", .y, "_sdq.sav")))
) |>
  set_names(paste0(combos$country, "_r", combos$round, "_sdq"))

# Note: Generally, the SDQ files for rounds 3-5 are named occh_older_child. The only exception is for Peru rounds 4 and 5, the files are named olderchild-SAQ. 
```

### Variable transformations

Key variables from each file

- Constructed files (ethiopia, india, peru, vietnam): child's ID (*childid*), cohort (*yc*), round, cluster ID (*clustid),* age in years (*ageyea*)*,* **sex** (*chsex*), child's ethnic group (*chethnic*), **child's religious affiliation** (*chldrel*), **life satisfaction** (*cladder*)

- SDQ-E files (ethiopia_r3-5, india_r3-5, peru_r3-5, vietnam_r3-5): strengths and difficulties questionnaire emotional subscale - **psychological distress** (comprised of 5 items) collected at rounds 3, 4 and 5 for the 1994-cohort at ages 15, 19, and 22 respectively

#### Select variables

*Constructed file*

```{r main-file-variables}
#| code-fold: true

# Variables from constructed file
keep_vars <- c("childid", "yc", "round", "ageyea", "chldrel", "cladder", "chsex", "chethnic", "clustid")
data <- map(data_list, function(df) {
  df |> 
    select(all_of(keep_vars)) |> 
    mutate(childid = as.character(childid))
})
```

*SDQ-E files*

```{r sdq-variables}
#| code-fold: true

## Country prefixes for childid
prefixes <- c(ethiopia = "ET", india = "IN", peru = "PE", vietnam = "VN")

## Fix childid: make sure there are 6 digits and add country prefix
fix_childid <- function(id, prefix) {
  id <- as.character(id)
  ifelse(str_detect(id, "^[A-Za-z]"),                 # if already full format
         id,                                          # leave it alone
         paste0(prefix, str_pad(id, 6, "left", "0"))) # otherwise reconstruct
}

## SDQ items come in two naming schemes across countries/rounds
scheme_headac <- function(r) c(
  headache = paste0("HEADACR", r),
  worry    = paste0("WRYLOTR", r),
  unhappy  = paste0("UNHPPYR", r),
  nervous  = paste0("NRVSITR", r),
  fears    = paste0("MNYFERR", r)
)
scheme_feay <- function(r) c(
  headache = paste0("FEAY13R", r),
  worry    = paste0("FEAY05R", r),
  unhappy  = paste0("FEAY20R", r),
  nervous  = paste0("FEAY38R", r),
  fears    = paste0("FEAY28R", r)
)

## Which scheme each country uses at R4/R5 (R3 is same across countries)
country_scheme <- list(
  india    = scheme_feay,
  peru     = scheme_headac,
  ethiopia = scheme_feay,
  vietnam  = scheme_feay
)

## Clean SDQ files: select + rename items and ID, then fix childid
sdq <- imap(sdq_list, function(df, nm) {
  ctry <- str_extract(nm, "^[a-z]+")
  rnum <- str_extract(nm, "[3-5]")
  pfx  <- prefixes[[ctry]]
  
  df <- df |> rename_with(toupper)
  
  # R3 uses HEADAC everywhere; R4/R5 use the country's scheme
  items <- if (rnum == "3") scheme_headac(rnum) else country_scheme[[ctry]](rnum)
  items <- setNames(items, paste0(names(items), "_r", rnum))
  
  idc <- if (rnum == "3") "CHILDID" else "CHILDCODE"
  
  df |>
    select(all_of(c(childid = idc, items))) |>
    mutate(childid = fix_childid(childid, pfx))
})
```

#### Check and transform SDQ

*Check values*

```{r inspect-sdq}
#| code-fold: true

## Distinct response values in each country-round, pre-rescale
imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  sort(unique(as.numeric(unlist(df[cols]))))
})
```

#### Test measurement invariance

The problem is that across rounds within the same country they use different response options. 
```{r measurement-invariance}
```

-------------------

SDQ-E needs to be **rescaled and standardised** as measurement was different across rounds and countries.

- Round 3: 1 - not true for you, 2 - a little true for you, 3 - certainly true for you, 77 - NK, 88 - N/A, 99 - Missing

- Rounds 4 and 5: 1 - strongly disagree, 2 - disagree, 3 - agree, 4 - strongly agree, 77 - NK, 79 - refused to answer

SDQ sub-scale is normally **0 - not true, 1 - somewhat true, 2 - certainly true -\> 10 full score**.

Therefore, scale of 1-3 (round 3) will be rescaled to 0-2. As for scale of 1-4 (rounds 4-5), 1,2 will be collapsed to 0 (not true), 3\>1, and 4-\>2.

The full score will be calculated as the mean of non-missing data \* 5.

*Rescale SDQ*

```{r rescale-sdq}

# Rescale function; unrecognised values (77/79/88/99) fall all become NA
rescale_3pt <- function(x) case_when(x == 1 ~ 0, x == 2 ~ 1, x == 3 ~ 2, TRUE ~ NA_real_)
rescale_4pt <- function(x) case_when(x %in% c(1, 2) ~ 0, x == 3 ~ 1, x == 4 ~ 2, TRUE ~ NA_real_)

# Format at R4/R5 (R3 is 3pt for everyone) — all four confirmed against the data
format_r45 <- c(india = "4pt", peru = "3pt", ethiopia = "4pt", vietnam = "4pt")

sdq <- imap(sdq, function(df, nm) {
  ctry <- str_extract(nm, "^[a-z]+")
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  
  fmt     <- if (rnum == "3") "3pt" else format_r45[[ctry]]
  rescale <- if (fmt == "3pt") rescale_3pt else rescale_4pt
  
  df |>
    mutate(across(all_of(cols), ~ rescale(as.numeric(.x)))) |>
    mutate(
      nonmiss = rowSums(!is.na(across(all_of(cols)))),
      !!paste0("emo_r", rnum) := if_else(
        nonmiss > 2, # need to have answered 3 or more out of 5 items to be included
        round(rowMeans(across(all_of(cols)), na.rm = TRUE) * 5),
        NA_real_
      )
    ) |>
    select(-nonmiss)
})
```

*Check rescaled values*

```{r check-sdq}
## Check each round's total emotional score is within 0-10 with sensible NA counts
imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  summary(df[[paste0("emo_r", rnum)]])
})
```

#### Transform variables

*Check values of variables - life satisfaction, sex, religion, ethnicity/caste, age*

```{r check-variables}
#| include: false

# Check distinct values of each key variable 
map(data, function(df) {
  map(df[c("cladder", "chsex", "chldrel", "chethnic", "ageyea")], ~sort(unique(as.numeric(.x))))
})
# No NA values 77/88/99 to transform

```

```{r variables-named}
#| code-fold: true 

# Transform into named categories
map(data, ~ haven::as_factor(.x$chldrel) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chsex) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chethnic) |> table(useNA = "always"))
```

*Create separate dataframes for each country*

```{r country-dataframes}
#| code-fold: true

# SDQ: one row per child 
sdq_scores <- imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  df |> select(childid, !!paste0("emo_r", rnum))
})

# Merge constructed data and SDQ for each country 
countries <- c("ethiopia", "india", "peru", "vietnam")

merged_data <- map(countries, function(ctry) {
  base      <- data[[ctry]]
  sdq_files <- sdq_scores[str_starts(names(sdq_scores), paste0(ctry, "_"))]
  reduce(sdq_files, left_join, by = "childid", .init = base)
}) |>
  set_names(countries)

# Row counts shouldn't jump (a jump = duplicate childids somewhere)
map_int(merged_data, nrow)

# Create separate dataframes for each country
list2env(merged_data, envir = .GlobalEnv) 

# For each child, take round 1's clustid for clustering
r1_clust <- map(merged_data, ~ .x |> filter(round ==1) |> 
                  mutate(clustid = if_else(clustid %in% c(0,80,88,90,99), NA_real_, as.numeric(clustid))) |> 
                  select(childid, clustid_base = clustid))
```

### Structuring data

#### Create variable 'equivalent age'

Although we have age in years as the variable for real age, we need an equivalent age of each round i.e. in round 3 for example, the 2001-cohort is aged 8 and the 1994-cohort is aged 15.

```{r equivalent-age}
#| code-fold: true

age_yc <- c("1"=1, "2"=5, "3"=8, "4"=12, "5"=15, "6"=19, "7"=22)
age_oc <- c("1"=8, "2"=12, "3"=15, "4"=19, "5"=22, "6"=26, "7"=29)

add_age <- function(df) {
  df %>%
    mutate(equivalent_age = if_else(yc == 1, age_yc[as.character(round)],
                                    age_oc[as.character(round)])) %>%
    filter(!is.na(equivalent_age)) %>%
    group_by(childid, equivalent_age) %>%
    slice_max(!is.na(cladder), n = 1, with_ties = FALSE) %>% # In round 6 there are multiple rows for phone calls but we only keep the row that has cladder data
    ungroup()
}

stacked <- map(merged_data, add_age)

stacked <- map(stacked, \(df) {
  df |>
    mutate(sdq = dplyr::case_when(
      round == 3 ~ emo_r3,
      round == 4 ~ emo_r4,
      round == 5 ~ emo_r5,
      TRUE ~ NA_real_
    )) |>
    select(-emo_r3, -emo_r4, -emo_r5)
})
names(stacked$ethiopia)   # emo_r3-5 combined into sdq
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Categorise minority and majority status

::: callout-note
**Update**: We will take into account country context when classifying majority/minority rather than solely rely on World Religion Database.
:::

World Religion Database religious demographics from 2000:

- Ethiopia: 58% Christian, 34% Muslim, 7.0% Ethnic religionists

- India: 75% Hindus, 13% Muslims, 4.2% Christians, 3.3% Ethnic religionists, 1.8% Sikhs, 1.4% Agnostics

- Peru: 96% Christian, 1.4% Ethnic religionist, 1.2% Agnostic

- Vietnam: 48% Buddhist, 13% Agnostic, 12% Ethnic religionist, 10.8% New religionist, 8.3% Christian, 6.7% Atheist, 1.0% Chinese folk religionist

*Check sample n by religion*

```{r check-religions}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chldrel) |>
      count(chldrel = as_factor(chldrel)))
```

For now, we will define the majority religions for each country as the following

- Ethiopia: **Orthodox**
- India: **Hinduism**
- Peru**: Catholicism**
- Vietnam: **Buddhism, none, and ancestor worship** (to be discussed more, but all of these could be considered dominant traditions in Vietnam)

*Classify majority/minority*

```{r classify}

# Specify majority religions
majority_religion <- list(ethiopia = "orthodox", india = "hindu", peru = "catholic", vietnam = c("buddhist", "none", "ancestor worship"))

classify <- function(df, ctry) {
  maj <- majority_religion[[tolower(ctry)]]
  df |> 
    mutate(
      chldrel = as.character(as_factor(chldrel)),
     minority = case_when(
       is.na(chldrel) ~ NA_character_,
       tolower(chldrel) %in% maj ~ "Majority",
       TRUE ~ "Minority"
     )
      )
}

stacked <- imap(stacked, classify)
```

As sensitivity check in later analyses, if there are variations across Young Lives regions, instead of country-level demographics we will use regional demographics to determine majority and minority status.

*Check classifications*

```{r check-classifications}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chldrel, minority) |>
      count(chldrel, minority))

map(stacked, ~ .x |>
      distinct(childid, chldrel, minority) |>
      count(minority) |>
      mutate(pct = round(100 * n / sum(n), 1)))

# Check how many NAs, then drop
stacked <- imap(stacked, classify) |> map(~ filter(.x, !is.na(minority)))
```

Missingness for religion (*chldrel*)

- Ethiopia: 2

- India: 0

- Peru: 1

- Vietnam: 1

------------------------------------------------------------------------

------------------------------------------------------------------------

# Descriptives

## Visualisations

### Life satisfaction - combined cohorts

*Overall trajectory*

```{r overall-traj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(4, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory)
wrap_plots(plots, ncol = 2)
```

*By cohort*

```{r traj-cohort}
#| code-fold: true

# Mean cladder by age and cohort, at overlap ages, per country
cohort_check <- map(stacked, ~ .x |>
                      filter(!is.na(cladder), equivalent_age %in% c(8, 12, 15, 19, 22)) |>
                      mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
                      group_by(equivalent_age, cohort) |>
                      summarise(mean_cl = mean(cladder), se = sd(cladder)/sqrt(n()), n = n(),
                                .groups = "drop"))

cohort_check$ethiopia

# Plot 
bind_rows(stacked, .id = "country") |>
  filter(!is.na(cladder), equivalent_age %in% c(8, 12, 15, 19, 22)) |>
  mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
  group_by(country, equivalent_age, cohort) |>
  summarise(mean_cl = mean(cladder), se = sd(cladder)/sqrt(n()), .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = cohort, fill = cohort)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  facet_wrap(~ country) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22)) +
  labs(y = "Mean life satisfaction", x = "Age") +
  theme_minimal()
```

***By minority/majority status***

```{r traj-min-maj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_minmaj <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age, minority) |>
    summarise(mean_cl = mean(cladder),
              se = sd(cladder)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(4, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = minority), alpha = 0.2, colour = NA) +
    scale_colour_manual(values = group_cols) +
    scale_fill_manual(values = group_cols) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_minmaj)
wrap_plots(plots, ncol = 2)
```

*By sex*

```{r traj-sex}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_sex <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    mutate(chsex = as_factor(chsex)) |> 
    group_by(equivalent_age, chsex) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = chsex)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(3, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sex)
wrap_plots(plots, ncol = 2)
```

*By majority/minority status and sex*

```{r traj-sex-minority}
#| code-fold: true

bind_rows(stacked, .id = "country") |>
  filter(!is.na(cladder), !is.na(equivalent_age)) |>
  mutate(chsex = as_factor(chsex)) |>
  group_by(country, equivalent_age, chsex, minority) |>
  summarise(mean_cl = mean(cladder),
            se = sd(cladder) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = minority, fill = minority)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96 * se, ymax = mean_cl + 1.96 * se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  scale_colour_manual(values = group_cols) +
  scale_fill_manual(values = group_cols) +
  facet_grid(country ~ chsex) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  scale_y_continuous(limits = c(4, 7)) +
  theme_minimal()
```

------------------------------------------------------------------------

### Psychological distress (SDQ-E) - 1994-cohort only

*Overall trajectory*

```{r overall-traj-sdq}
#| code-fold: true

# Mean SDQ-E by age, overall and by group, per country
plot_trajectory_sdq <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(sdq), !is.na(equivalent_age)) |>
    group_by(equivalent_age) |>
    summarise(mean_cl = mean(sdq), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(title = ctry, x = "Age", y = "Mean psychological distress") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sdq)
wrap_plots(plots, ncol = 2)
```

***By minority/majority status***

```{r traj-min-maj-sdq}
#| code-fold: true

# Mean SDQ by age, overall and by group, per country
plot_trajectory_minmaj_sdq <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(sdq), !is.na(equivalent_age)) |>
    group_by(equivalent_age, minority) |>
    summarise(mean_cl = mean(sdq),
              se = sd(sdq)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(title = ctry, x = "Age", y = "Mean psychological distress") +
    scale_colour_manual(values = group_cols) +
    scale_fill_manual(values = group_cols) +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = minority), alpha = 0.2, colour = NA) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_minmaj_sdq)
wrap_plots(plots, ncol = 2)
```

*By sex*

```{r traj-sex-sdq}
#| code-fold: true

# Mean SDQ by age, overall and by group, per country
plot_trajectory_sex_sdq <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(sdq), !is.na(equivalent_age)) |>
    mutate(chsex = as_factor(chsex)) |> 
    group_by(equivalent_age, chsex) |>
    summarise(mean_cl = mean(sdq), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = chsex)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(0, 10)) +
    labs(title = ctry, x = "Age", y = "Mean psychological distress") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sex_sdq)
wrap_plots(plots, ncol = 2)
```

*By majority/minority status and sex*

```{r traj-sex-minority-sdq}
#| code-fold: true

bind_rows(stacked, .id = "country") |>
  filter(!is.na(sdq), !is.na(equivalent_age)) |>
  mutate(chsex = as_factor(chsex)) |>
  group_by(country, equivalent_age, chsex, minority) |>
  summarise(mean_cl = mean(sdq),
            se = sd(sdq) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = minority, fill = minority)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96 * se, ymax = mean_cl + 1.96 * se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  facet_grid(country ~ chsex) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  scale_y_continuous(limits = c(0, 10)) +
  scale_colour_manual(values = group_cols) +
  scale_fill_manual(values = group_cols) +
  theme_minimal()
```

------------------------------------------------------------------------

## Descriptive statistics

*Summary table: per country x age x religious majority/minority*

```{r table-descriptives}
#| code-fold: true

# All countries together
all_df <- bind_rows(stacked, .id = "country")

# Pivot longer, mean/SD/n per country x age x minority group
desc <- all_df |> 
  pivot_longer(c(cladder, sdq), names_to = "outcome", values_to = "value") |> 
  filter(!is.na(value), !is.na(minority)) |> 
  group_by(country, outcome, equivalent_age, minority) |> 
  summarise(mean = mean(value), sd = sd(value), n = n(), .groups = "drop")

# Clear wide table - maj/min side by side
desc_table <- desc |> 
  mutate(cell = sprintf("%.2f (%.2f) [n=%d]", mean, sd, n)) |> 
  select(country, outcome, equivalent_age, minority, cell) |> 
  pivot_wider(names_from = minority, values_from = cell) |> 
  arrange(country, outcome, equivalent_age)

desc_table
```

*Group comparisons at each timepoint - Welch's t-test*

```{r comparisons}
#| code-fold: true

comparisons <- all_df |> 
  pivot_longer(c(cladder, sdq), names_to = "outcome", values_to = "value") |> 
  filter(!is.na(value), !is.na(minority)) |> 
  mutate(minority = factor(minority, levels = c("Minority", "Majority"))) |> 
  group_by(country, outcome, equivalent_age) |> 
  filter(n_distinct(minority) == 2,
         min(table(minority)) >= 2) |> 
  summarise(
    n_min = sum(minority == "Minority"),
    n_maj = sum(minority == "Majority"),
    tt = list(t.test(value ~ minority)),
    .groups = "drop"
  ) |> 
  mutate(
    diff_min_maj = map_dbl(tt, ~ .x$estimate[1] - .x$estimate[2]),
    ci_low = map_dbl(tt, ~ .x$conf.int[1]),
    ci_high = map_dbl(tt, ~ .x$conf.int[2]),
    p_value = map_dbl(tt, "p.value")
  ) |> 
  select(-tt)
comparisons
```

Welch's t test is preferred since the minority and majority groups do not have equal variances nor sample sizes. We can test whether the group means between minorities and majorities are statistically significant.

Note that these are purely descriptive, we're doing many independent t-tests with no correction so this is more for establishing overall patterns.

*diff_min_maj* is the minority - majority, i.e. if *diff_min_maj* is positive for *cladder*, that means that minorities have higher life satisfaction; whereas if it's positive for *sdq*, that means minorities have lower psychological distress (since the two scales are coded differently).

Summary

- Ethiopia: cladder at age 15, 22, 26, 29: minority better life satisfaction than majority

- India: cladder at age 22: minority better life satisfaction than majority

------------------------------------------------------------------------

## Variability

```{r histograms}
#| code-fold: true

plot_hist <- function(df, ctry, outcome_var) {
  df |> 
    filter(!is.na(.data[[outcome_var]]), !is.na(minority)) |> 
    ggplot(aes(x = .data[[outcome_var]], fill = minority)) +
    geom_histogram(aes(y = after_stat(density)),
                   binwidth = 1, alpha = 0.5, position = "identity") +
    facet_wrap(~ equivalent_age, ncol = 4) +
    labs(title = paste0(ctry, " - ", outcome_var), x = outcome_var, y = "Count") +
    theme_minimal() +
    scale_fill_manual(values = group_cols) +
    theme(legend.position = "none")
}

# Life satisfaction (combined cohort: 7 ages x 2 groups)
iwalk(stacked, ~ print(plot_hist(.x, .y, "cladder")))

# Psychological distress (1994-cohort: 3 ages x 2 groups)
iwalk(stacked, ~ print(plot_hist(.x, .y, "sdq")))
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Latent Growth Curve Analyses (LGCA)

*Structure data*

```{r pivot-wide}
#| code-fold: true

to_wide <- function(df) {
  df |>
    mutate(chsex = as.numeric(as_factor(chsex)) - 1,  # 0/1
           minority = if_else(minority == "Minority", 1, 0)) |>
    select(childid, chsex, minority, clustid, yc, equivalent_age, cladder, sdq) |>
    pivot_wider(names_from = equivalent_age,
                values_from = c(cladder, sdq),
                names_prefix = "age")
}
wide <- map(stacked, to_wide)
wide <- map(wide, ~ .x |> group_by(childid) |>
              fill(clustid, .direction = "downup") |> ungroup())

names(wide$ethiopia)  # check column names
```

*Filter children* (only those with 3 complete cladder out of 5 for 2001-cohort, 7 for 1994-cohort)

```{r filter-sample}
#| code-fold: true

wide <- map(wide, ~ .x |> 
              mutate(n_cl = rowSums(!is.na(across(starts_with("cladder_age"))))) |> 
              filter(n_cl >= 3) |> 
              select(-n_cl))
```

## Life satisfaction models

Given the two cohorts are not equivalent - with diverging trajectories, we will fit the models separately, and religious minority status and sex predicting the growth factors. 2001-cohort will have age 8-22, and 1994-cohort 12-29.

We have two cohorts across four countries, so for simplicity we will build a function so it can be easily replicated.

*LGCA helpers*

```{r lgca-helpers}
#| code-fold: true

# Build lavaan growth syntax from a vector of ages, centred at the first age.
# type: "intercept", "linear", or "conditional"
make_spec <- function(ages, shape = "linear", conditional = FALSE, extra = NULL) {
  t    <- (ages - ages[1]) / 10          # decades 
  cols <- paste0("cladder_age", ages)
  i    <- paste0("i =~ ", paste0("1*", cols, collapse = " + "))

  if (shape == "intercept") {
    spec <- i; factors <- "i"

  } else if (shape == "basis") {                       # latent-basis (freed-loading)
    load <- rep("NA", length(ages))                    # middle loadings free
    load[1] <- "0"; load[length(ages)] <- "1"          # first fixed 0, last fixed 1
    s <- paste0("s =~ ", paste0(load, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")

  } else {
    s <- paste0("s =~ ", paste0(t, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")
    if (shape %in% c("quadratic", "cubic")) {
      q <- paste0("q =~ ", paste0(t^2, "*", cols, collapse = " + "))
      spec <- paste(spec, q, sep = "\n"); factors <- c(factors, "q")
    }
    if (shape == "cubic") {
      cub <- paste0("cub =~ ", paste0(t^3, "*", cols, collapse = " + "))
      spec <- paste(spec, cub, sep = "\n"); factors <- c(factors, "cub")
    }
  }

  if (conditional) {
    spec <- paste(c(spec, paste0(factors, " ~ minority + chsex")), collapse = "\n")
  }
  if(!is.null(extra)){
    spec <- paste(c(spec, extra), collapse ="\n")
  }
  spec
}

# Ages available per cohort per country
ages_yc <- list(ethiopia = c(8,12,15,19,22), india = c(8,12,15,19,22),
                peru = c(8,12,15,19,22), vietnam = c(8,12,15,19)) # vietnam up to age 19 only
ages_oc <- list(ethiopia = c(12,15,19,22,26,29), india = c(12,15,19,22,26,29),
                peru = c(12,15,19,22,26,29), vietnam = c(12,15,19,22,26)) # vietnam up to age 26 only

# Fit one model type for one cohort, across all countries
fit_all <- function(ages_list, shape, cohort_val, conditional = FALSE, constraints = NULL) {
  imap(wide, function(df, ctry) {
    spec <- make_spec(ages_list[[ctry]], shape, conditional = conditional, extra = constraints[[ctry]])
    d <- df |> filter(yc == cohort_val, !is.na(minority), !is.na(chsex)) |>
      mutate(minority = as.numeric(minority), chsex = as.numeric(chsex))
    tryCatch(growth(spec, data = d, missing = "fiml"), error = function(e) NULL)
  })
}

get_fit <- function(fits) {
  extract <- function(f) {
    if (is.null(f)) return(NULL)
    if (!lavInspect(f, "converged")) return(NULL)   # skip non-converged
    as_tibble(as.list(fitMeasures(f, c("cfi","tli","rmsea"))))
  }
  map_dfr(fits, extract, .id = "country")
}

get_effects <- function(fits) {
  extract <- function(f) {
    if (is.null(f)) return(NULL)
    if (!lavInspect(f, "converged")) return(NULL)
    parameterEstimates(f) |>
      filter(rhs %in% c("minority", "chsex"),
             lhs %in% c("i", "s", "q", "cub")) |>  
      select(lhs, rhs, est, se, pvalue)
  }
  map_dfr(fits, extract, .id = "country")
}

check_model <- function(fit) {
  if (is.null(fit)) return("NULL (errored)")
  if (!lavInspect(fit, "converged")) return("did not converge")
  pe   <- parameterEstimates(fit)
  vars <- pe[pe$op == "~~" & pe$lhs == pe$rhs, ]     # all variances
  if (any(vars$est < 0)) return("Heywood (negative variance)")
  "OK"
}
```

### 2001-cohort age 8-22

*Intercept only*

```{r yc-intercept}
yc_int <- fit_all(ages_yc, "intercept", 1)
get_fit(yc_int)
```

*Linear (intercept and slope), FIML with clustering by clustid*

```{r yc-linear}
yc_lin <- fit_all(ages_yc, "linear", 1)
get_fit(yc_lin)
```

*Conditional model (minority and sex)*

```{r yc-conditional}
yc_cond <- fit_all(ages_yc, "linear", 1, conditional = TRUE)
get_effects(yc_cond)
```

*Quadratic*

```{r quadratic}
yc_quad <- fit_all(ages_yc, "quadratic", 1)
get_fit(yc_quad)

# No cubic for 2001-cohort; not enough time-points
```
*Conditional, quadratic*
```{r yc-conditional-quadratic}
yc_cond <- fit_all(ages_yc, "quadratic", 1, conditional = TRUE)
get_effects(yc_cond)
```

*Basis model*
```{r yc-basis}
yc_basis <- fit_all(ages_yc, "basis", 1,
                    constraints = list(ethiopia = "s ~~ 0*s")) # specify ethiopia constraints separately
purrr::map_chr(yc_basis, check_model)     
get_fit(yc_basis)

yc_basis_cond <- fit_all(ages_yc, "basis", 1, conditional = TRUE,
                   constraints = list(ethiopia = "s ~~ 0*s"))
get_effects(yc_basis_cond)
```

We use basis model / freed-loading model because trajectories were not monotonic - it seemed to decline throughout adolescence but partially recover over time. Linear and quadratic models fit poorly, but latent-basis (freed-loading) growth model can estimate the shape of average trajectory from the data rather than imposing a functional form.

In basis models, the first and last factor loadings are fixed to 0 and 1 respectively and in between loadings are freely estimated. Intercept = expected outcome at initial stage; slope = total change across observed age range; estimated loadings = proportion of total change realised at each save. 

Model fit is now substantially improved with the freed-loading specification. 

In Ethiopia 2001-cohort, the slope-factor variance was non-significant and marginally negative (Heywood case), suggesting negligible variability in trajectory of change between individauls. To fit a model we thus have to constrain the slope-factor variance to 0. This only affects estimation of individual differences in slope, not the mean trajectory or estimated minority-majority effects.

------------------------------------------------------------------------

### 1994-cohort age 12-29

*Intercept-only*

```{r oc-intercept}
oc_int <- fit_all(ages_oc, "intercept", 0)
get_fit(oc_int)
```

*Linear (intercept + slope)*

```{r oc-linear}
oc_lin <- fit_all(ages_oc, "linear", 0)
get_fit(oc_lin)
```

*Conditional (minority + sex)*

```{r oc-conditional}
oc_cond <- fit_all(ages_oc, "linear", 0, conditional = TRUE)
get_effects(oc_cond)
```

*Quadratic and cubic*

```{r quad-cub}
oc_quad <- fit_all(ages_oc, "quadratic", 0)
oc_cubic <- fit_all(ages_oc, "cubic", 0)   

get_fit(oc_quad)
get_fit(oc_cubic)

# It seems cubic failed to converge for all countries
```

*Conditional, quadratic*
```{r oc-conditional}
oc_cond <- fit_all(ages_oc, "quadratic", 0, conditional = TRUE)
get_effects(oc_cond)
```

*Basis model*
```{r oc-basis}
oc_basis <- fit_all(ages_oc, "basis", 0) 
purrr::map_chr(oc_basis, check_model) 
get_fit(oc_basis)

oc_basis_cond <- fit_all(ages_oc, "basis", 0, conditional = TRUE)
get_effects(oc_basis_cond)
```

### Check fit indices

```{r fit-indices}
bind_rows(
  get_fit(yc_lin)  |> mutate(cohort = "Younger", model = "Linear"),
  get_fit(oc_lin)  |> mutate(cohort = "Older",   model = "Linear"),
  get_fit(yc_quad) |> mutate(cohort = "Younger", model = "Quadratic"),
  get_fit(oc_quad) |> mutate(cohort = "Older",   model = "Quadratic"),
  get_fit(yc_basis) |> mutate(cohort = "Younger", model = "Basis"),
  get_fit(oc_basis) |> mutate(cohort = "Older", model = "Basis")
) |>
  relocate(country, cohort, model) |>
  mutate(across(c(cfi, tli, rmsea), ~ round(.x, 3)))
```

Fit indices are... poor. CFI and TLI are not \>=0.95, RMSEA mostly not \<=0.06 (except Vietnam).

Even for linear the fit is poor, and quadratic does not do any better. Keep linear for parsimony.

------------------------------------------------------------------------

### Trajectory visualisations

Visualise model-implied trajectories of minority vs majority groups per country

#### Linear

```{r build-plot-traj}
#| code-fold: true 

# Build implied minority & majority trajectories from a conditional fit
implied_traj <- function(fit, ages) {
  if (is.null(fit)) return(NULL)
  cf <- coef(fit)
  t <- ages - ages[1]
  # majority = reference (minority=0, sex held at 0); minority adds its coef
  maj_i <- cf["i~1"]; maj_s <- cf["s~1"]
  min_i <- maj_i + cf["i~minority"]; min_s <- maj_s + cf["s~minority"]
  tibble(
    age = rep(ages, 2),
    group = rep(c("Majority","Minority"), each = length(ages)),
    fit = c(maj_i + maj_s*t, min_i + min_s*t)
  )
}
```

```{r plot-traj}
# combine observed means + model lines, one plot per cohort
plot_obs_model <- function(fits, ages_list, cohort_val, title) {
  # model-implied lines
  model_df <- imap_dfr(fits, function(f, ctry) {
    d <- implied_traj(f, ages_list[[ctry]])
    if (is.null(d)) return(NULL)
    mutate(d, country = ctry)
  })
  # observed means
  obs_df <- imap_dfr(stacked, function(df, ctry) {
    df |> filter(yc == cohort_val, !is.na(cladder),
                 equivalent_age %in% ages_list[[ctry]]) |>
      mutate(group = if_else(minority == "Minority", "Minority", "Majority")) |>
      group_by(country = ctry, age = equivalent_age, group) |>
      summarise(mean_cl = mean(cladder), .groups = "drop")
  })
  
  ggplot() +
    geom_point(data = obs_df, aes(age, mean_cl, colour = group), alpha = 0.6) +
    geom_line(data = model_df, aes(age, fit, colour = group), linewidth = 1) +
    facet_wrap(~ country) +
    scale_colour_manual(values = c("Majority"="#4477AA","Minority"="#CC6677")) +
    labs(title = title, y = "Life satisfaction", x = "Age", colour = NULL) +
    theme_minimal()
}

plot_obs_model(yc_cond, ages_yc, 1, "2001-Cohort")
plot_obs_model(oc_cond, ages_oc, 0, "1994-Cohort")
```

#### Basis 

```{r traj-basis}
implied_traj_basis <- function(fit, ages) {
  if(is.null(fit)) return(NULL)
  if(!lavInspect(fit, "converged")) return(NULL)
  
  est <- parameterEstimates(fit)
  cols <- paste0("cladder_age", ages)
  
  # estimated basis loadings on s (freed loadings), pulled in wave order
  load <- sapply(cols, function(cc)
    est$est[est$op == "=~" & est$lhs == "s" & est$rhs == cc])
  
  cf <- coef(fit)
  maj_i <- cf["i~1"]; maj_s <- cf["s~1"]
  min_i <- maj_i + cf["i~minority"]
  min_s <- maj_s + cf["s~minority"]
  
  tibble(
    age = rep(ages, 2),
    group = rep(c("Majority", "Minority"), each = length(ages)),
    fit = c(maj_i + load * maj_s, # majority: intercept + loading x slope
            min_i + load * min_s) # minority: same, shifted by minority coefs
  )
}
```

```{r plot-basis}
plot_obs_model <- function(fits, ages_list, cohort_val, title) {
  model_df <- imap_dfr(fits, function(f, ctry) {
    d <- implied_traj_basis(f, ages_list[[ctry]])
    if (is.null(d)) return(NULL)
    mutate(d, country = ctry)
  })

  obs_df <- imap_dfr(stacked, function(df, ctry) {
    df |> filter(yc == cohort_val, !is.na(cladder),
                 equivalent_age %in% ages_list[[ctry]]) |>
      mutate(group = if_else(minority == "Minority", "Minority", "Majority")) |>
      group_by(country = ctry, age = equivalent_age, group) |>
      summarise(mean_cl = mean(cladder),
                se = sd(cladder) / sqrt(n()),          # <- SE of the mean
                .groups = "drop") |>
      mutate(lo = mean_cl - 1.96 * se,
             hi = mean_cl + 1.96 * se)
  })

  ggplot() +
    geom_ribbon(data = obs_df,
                aes(age, ymin = lo, ymax = hi, fill = group),
                alpha = 0.15, colour = NA) +               # <- the bands
    geom_point(data = obs_df, aes(age, mean_cl, colour = group), alpha = 0.6) +
    geom_line(data = model_df, aes(age, fit, colour = group), linewidth = 1) +
    facet_wrap(~ country) +
    scale_colour_manual(values = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
    scale_fill_manual(values   = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
    labs(title = title, y = "Life satisfaction", x = "Age", colour = NULL, fill = NULL) +
    theme_minimal()
}

plot_obs_model(yc_basis_cond, ages_yc, 1, "2001-Cohort")
plot_obs_model(oc_basis_cond, ages_oc, 0, "1994-Cohort")
```


------------------------------------------------------------------------
## Psychological distress models 

*LGCA helpers*

Only intercept, linear and conditional - no higher order.

```{r lgca-helpers-sdq}
#| code-fold: true

# Build lavaan growth syntax from a vector of ages, centred at the first age.
# type: "intercept", "linear", or "conditional"
make_spec_sdq <- function(ages, shape = "linear", conditional = FALSE, extra = NULL) {
  t    <- (ages - ages[1]) / 10          # decades 
  cols <- paste0("sdq_age", ages)
  i    <- paste0("i =~ ", paste0("1*", cols, collapse = " + "))

  if (shape == "intercept") {
    spec <- i; factors <- "i"

  } else if (shape == "basis") {                       # latent-basis (freed-loading)
    load <- rep("NA", length(ages))                    # middle loadings free
    load[1] <- "0"; load[length(ages)] <- "1"          # first fixed 0, last fixed 1
    s <- paste0("s =~ ", paste0(load, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")

  } else {
    s <- paste0("s =~ ", paste0(t, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")
  }

  if (conditional) {
    spec <- paste(c(spec, paste0(factors, " ~ minority + chsex")), collapse = "\n")
  }
  if(!is.null(extra)){
    spec <- paste(c(spec, extra), collapse ="\n")
  }
  spec
}

# Ages available per country
ages_oc <- list(ethiopia = c(15,19,22), india = c(15,19,22),
                peru = c(15,19,22), vietnam = c(15,19,22)) 
```

### 1994-cohort age 15-22

*Intercept-only*

```{r oc-intercept-sdq}
oc_int_sdq <- fit_all(ages_oc, "intercept", 0)
get_fit(oc_int)
```

*Linear (intercept + slope)*

```{r oc-linear-sdq}
oc_lin <- fit_all(ages_oc, "linear", 0)
get_fit(oc_lin)
```


*Conditional (minority + sex)*

```{r oc-conditional-sdq}
oc_cond <- fit_all(ages_oc, "linear", 0, conditional = TRUE)
get_effects(oc_cond)
```
