---
title: "Mental health trajectories of religious majority and minority youth: data pre-processing"
format: 
  html: 
    toc: true
    toc-depth: 3
    number-sections: true
    df-print: paged
    embed-resources: true
editor: visual
execute:
  echo: true
  warning: false
  message: false
---

This Quarto document pre-processes data from the Young Lives project, accessed via the UK Data Service on 4th August, 2026.

# Set up

*Load packages*

```{r setup}
#| code-fold: true

# Install any packages that aren't there yet then load
pacman::p_load(
  tidyverse, # data wrangling
  here, # clean file paths
  lavaan, # latent growth curve models
  haven, 
  patchwork, 
  purrr,
  ggplot2
)

# Set colours
group_cols <- c(Majority = "#4A90D9",   # blue
                Minority = "#E05B5B")   # red

options(scipen = 999)
```

*Load data*

```{r load-data}
#| code-fold: true

# here() builds a path from project root, so it works regardless of machine or subfolder
countries <- c("ethiopia", "india", "peru", "vietnam")

# Load rounds 1-7 constructed files 
data_list <- map(countries, ~ read_sav(here("data", paste0(.x, ".sav")))) |>
  set_names(countries)

# Load SDQ-E files (rounds 3-5, 1994-cohort only)
rounds    <- 3:5
combos <- tidyr::expand_grid(country = countries, round = rounds) # every country–round combination
sdq_list <- map2(
  combos$country, combos$round,
  ~ read_sav(here("data", paste0(.x, "_r", .y, "_sdq.sav")))
) |>
  set_names(paste0(combos$country, "_r", combos$round, "_sdq"))

# Note: Generally, the SDQ files for rounds 3-5 are named occh_older_child. The only exception is for Peru rounds 4 and 5, the files are named olderchild-SAQ. 
```

### Variable transformations

Key variables from each file

- Constructed files (ethiopia, india, peru, vietnam): child's ID (*childid*), cohort (*yc*), round, cluster ID (*clustid),* age in years (*ageyea*)*,* **sex** (*chsex*), child's ethnic group (*chethnic*), **child's religious affiliation** (*chldrel*), **life satisfaction** (*cladder*)

- SDQ-E files (ethiopia_r3-5, india_r3-5, peru_r3-5, vietnam_r3-5): strengths and difficulties questionnaire emotional subscale - **psychological distress** (comprised of 5 items) collected at rounds 3, 4 and 5 for the 1994-cohort at ages 15, 19, and 22 respectively

#### Select variables

*Constructed file*

```{r main-file-variables}
#| code-fold: true

# Variables from constructed file
keep_vars <- c("childid", "yc", "round", "ageyea", "chldrel", "cladder", "chsex", "chethnic", "clustid")
data <- map(data_list, function(df) {
  df |> 
    select(all_of(keep_vars)) |> 
    mutate(childid = as.character(childid))
})
```

*SDQ-E files*

```{r sdq-variables}
#| code-fold: true

## Country prefixes for childid
prefixes <- c(ethiopia = "ET", india = "IN", peru = "PE", vietnam = "VN")

## Fix childid: make sure there are 6 digits and add country prefix
fix_childid <- function(id, prefix) {
  id <- as.character(id)
  ifelse(str_detect(id, "^[A-Za-z]"),                 # if already full format
         id,                                          # leave it alone
         paste0(prefix, str_pad(id, 6, "left", "0"))) # otherwise reconstruct
}

## SDQ items come in two naming schemes across countries/rounds
scheme_headac <- function(r) c(
  headache = paste0("HEADACR", r),
  worry    = paste0("WRYLOTR", r),
  unhappy  = paste0("UNHPPYR", r),
  nervous  = paste0("NRVSITR", r),
  fears    = paste0("MNYFERR", r)
)
scheme_feay <- function(r) c(
  headache = paste0("FEAY13R", r),
  worry    = paste0("FEAY05R", r),
  unhappy  = paste0("FEAY20R", r),
  nervous  = paste0("FEAY38R", r),
  fears    = paste0("FEAY28R", r)
)

## Which scheme each country uses at R4/R5 (R3 is same across countries)
country_scheme <- list(
  india    = scheme_feay,
  peru     = scheme_headac,
  ethiopia = scheme_feay,
  vietnam  = scheme_feay
)

## Clean SDQ files: select + rename items and ID, then fix childid
sdq <- imap(sdq_list, function(df, nm) {
  ctry <- str_extract(nm, "^[a-z]+")
  rnum <- str_extract(nm, "[3-5]")
  pfx  <- prefixes[[ctry]]
  
  df <- df |> rename_with(toupper)
  
  # R3 uses HEADAC everywhere; R4/R5 use the country's scheme
  items <- if (rnum == "3") scheme_headac(rnum) else country_scheme[[ctry]](rnum)
  items <- setNames(items, paste0(names(items), "_r", rnum))
  
  idc <- if (rnum == "3") "CHILDID" else "CHILDCODE"
  
  df |>
    select(all_of(c(childid = idc, items))) |>
    mutate(childid = fix_childid(childid, pfx))
})
```

#### Check SDQ

*Check values*

```{r inspect-sdq}
#| code-fold: true

# Remove missing data 
missing <- c(77,79,88,99)
sdq <- imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  df |> 
    mutate(across(all_of(cols),
                  ~ if_else(as.numeric(.x) %in% missing, NA_real_,
                            as.numeric(.x))))
})

## Distinct response values in each country-round
imap(sdq, function(df, nm) {
  rnum <- str_extract(nm, "[3-5]")
  cols <- paste0(c("headache", "worry", "unhappy", "nervous", "fears"), "_r", rnum)
  sort(unique(as.numeric(unlist(df[cols]))))
})
```

There are different response options depending on rounds - for India, Ethiopia, and Vietnam, round 3 has 3 response options but rounds 4-5 have 4. For Peru, there are 3 response options across time.

#### Transform variables

*Check values of variables - life satisfaction, sex, religion, ethnicity/caste, age*

```{r check-variables}
#| include: false

# Check distinct values of each key variable 
map(data, function(df) {
  map(df[c("cladder", "chsex", "chldrel", "chethnic", "ageyea")], ~sort(unique(as.numeric(.x))))
})
# No NA values 77/88/99 to transform

```

```{r variables-named}
#| code-fold: true 

# Transform into named categories
map(data, ~ haven::as_factor(.x$chldrel) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chsex) |> table(useNA = "always"))
map(data, ~ haven::as_factor(.x$chethnic) |> table(useNA = "always"))
```

*Create separate dataframes for each country*

```{r country-dataframes}
#| code-fold: true

sdq_items <- c("headache", "worry", "unhappy", "nervous", "fears")
# SDQ: keep individual items as well as total
sdq_scores <- imap(sdq, function(df, nm) {
  rnum  <- str_extract(nm, "[3-5]")
  items <- paste0(sdq_items, "_r", rnum)          # headache_r3, ... fears_r3
  total <- paste0("emo_r", rnum)                  # keep the total too 
  keep  <- intersect(c(items, total), names(df))  # whatever's present
  df |> select(childid, all_of(keep))
})

# Merge constructed data and SDQ for each country 
countries <- c("ethiopia", "india", "peru", "vietnam")
merged_data <- map(countries, function(ctry) {
  base      <- data[[ctry]]
  sdq_files <- sdq_scores[str_starts(names(sdq_scores), paste0(ctry, "_"))]
  reduce(sdq_files, left_join, by = "childid", .init = base)
}) |>
  set_names(countries)

# Row counts shouldn't jump (a jump = duplicate childids somewhere)
map_int(merged_data, nrow)

# Create separate dataframes for each country
list2env(merged_data, envir = .GlobalEnv) 

# For each child, take round 1's clustid for clustering
r1_clust <- map(merged_data, ~ .x |> filter(round ==1) |> 
                  mutate(clustid = if_else(clustid %in% c(0,80,88,90,99), NA_real_, as.numeric(clustid))) |> 
                  select(childid, clustid_base = clustid))
```

### Structuring data

#### Create variable 'equivalent age'

Although we have age in years as the variable for real age, we need an equivalent age of each round i.e. in round 3 for example, the 2001-cohort is aged 8 and the 1994-cohort is aged 15.

```{r equivalent-age}
#| code-fold: true

age_yc <- c("1"=1, "2"=5, "3"=8, "4"=12, "5"=15, "6"=19, "7"=22)
age_oc <- c("1"=8, "2"=12, "3"=15, "4"=19, "5"=22, "6"=26, "7"=29)

add_age <- function(df) {
  df %>%
    mutate(equivalent_age = if_else(yc == 1, age_yc[as.character(round)],
                                    age_oc[as.character(round)])) %>%
    filter(!is.na(equivalent_age)) %>%
    group_by(childid, equivalent_age) %>%
    slice_max(!is.na(cladder), n = 1, with_ties = FALSE) %>% # In round 6 there are multiple rows for phone calls but we only keep the row that has cladder data
    ungroup()
}

stacked <- map(merged_data, add_age)

stacked <- map(stacked, \(df) {
  df |> mutate(
    # pick the item set for whichever round this row belongs to
    sdq = dplyr::case_when(
      round == 3 ~ rowSums(across(all_of(paste0(sdq_items, "_r3"))), na.rm = FALSE),
      round == 4 ~ rowSums(across(all_of(paste0(sdq_items, "_r4"))), na.rm = FALSE),
      round == 5 ~ rowSums(across(all_of(paste0(sdq_items, "_r5"))), na.rm = FALSE),
      TRUE ~ NA_real_
    )
  )
})
names(stacked$ethiopia)   # emo_r3-5 combined into sdq
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Categorise minority and majority status

::: callout-note
**Update**: We will take into account country context when classifying majority/minority rather than solely rely on World Religion Database.
:::

World Religion Database religious demographics from 2000:

- Ethiopia: 58% Christian, 34% Muslim, 7.0% Ethnic religionists

- India: 75% Hindus, 13% Muslims, 4.2% Christians, 3.3% Ethnic religionists, 1.8% Sikhs, 1.4% Agnostics

- Peru: 96% Christian, 1.4% Ethnic religionist, 1.2% Agnostic

- Vietnam: 48% Buddhist, 13% Agnostic, 12% Ethnic religionist, 10.8% New religionist, 8.3% Christian, 6.7% Atheist, 1.0% Chinese folk religionist

*Check sample n by religion*

```{r check-religions}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chldrel) |>
      count(chldrel = as_factor(chldrel)))
```

For now, we will define the majority religions for each country as the following

- Ethiopia: **Orthodox**
- India: **Hinduism**
- Peru**: Catholicism**
- Vietnam: **Buddhism, none, and ancestor worship** (to be discussed more, but all of these could be considered dominant traditions in Vietnam)

*Classify majority/minority*

```{r classify}

# Specify majority religions
majority_religion <- list(ethiopia = "orthodox", india = "hindu", peru = "catholic", vietnam = c("buddhist", "none", "ancestor worship"))

classify <- function(df, ctry) {
  maj <- majority_religion[[tolower(ctry)]]
  df |> 
    mutate(
      chldrel = as.character(as_factor(chldrel)),
     minority = case_when(
       is.na(chldrel) ~ NA_character_,
       tolower(chldrel) %in% maj ~ "Majority",
       TRUE ~ "Minority"
     )
      )
}

stacked <- imap(stacked, classify)
```

As sensitivity check in later analyses, if there are variations across Young Lives regions, instead of country-level demographics we will use regional demographics to determine majority and minority status.

*Check classifications*

```{r check-classifications}
#| code-fold: true

map(stacked, ~ .x |>
      distinct(childid, chldrel, minority) |>
      count(chldrel, minority))

map(stacked, ~ .x |>
      distinct(childid, chldrel, minority) |>
      count(minority) |>
      mutate(pct = round(100 * n / sum(n), 1)))

# Check how many NAs, then drop
stacked <- imap(stacked, classify) |> map(~ filter(.x, !is.na(minority)))
```

Missingness for religion (*chldrel*)

- Ethiopia: 2

- India: 0

- Peru: 1

- Vietnam: 1

------------------------------------------------------------------------

------------------------------------------------------------------------

# Descriptives

## Visualisations

### Life satisfaction - combined cohorts

*Overall trajectory*

```{r overall-traj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(4, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory)
wrap_plots(plots, ncol = 2)
```

*By cohort*

```{r traj-cohort}
#| code-fold: true

# Mean cladder by age and cohort, at overlap ages, per country
cohort_check <- map(stacked, ~ .x |>
                      filter(!is.na(cladder), equivalent_age %in% c(8, 12, 15, 19, 22)) |>
                      mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
                      group_by(equivalent_age, cohort) |>
                      summarise(mean_cl = mean(cladder), se = sd(cladder)/sqrt(n()), n = n(),
                                .groups = "drop"))

cohort_check$ethiopia

# Plot 
bind_rows(stacked, .id = "country") |>
  filter(!is.na(cladder), equivalent_age %in% c(8, 12, 15, 19, 22)) |>
  mutate(cohort = if_else(yc == 1, "Younger", "Older")) |>
  group_by(country, equivalent_age, cohort) |>
  summarise(mean_cl = mean(cladder), se = sd(cladder)/sqrt(n()), .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = cohort, fill = cohort)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  facet_wrap(~ country) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22)) +
  labs(y = "Mean life satisfaction", x = "Age") +
  theme_minimal()
```

***By minority/majority status***

```{r traj-min-maj}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_minmaj <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    group_by(equivalent_age, minority) |>
    summarise(mean_cl = mean(cladder),
              se = sd(cladder)/sqrt(n()),
              n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = minority)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(4, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    geom_ribbon(aes(ymin = mean_cl - 1.96*se, ymax = mean_cl + 1.96*se, fill = minority), alpha = 0.2, colour = NA) +
    scale_colour_manual(values = group_cols) +
    scale_fill_manual(values = group_cols) +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_minmaj)
wrap_plots(plots, ncol = 2)
```

*By sex*

```{r traj-sex}
#| code-fold: true

# Mean cladder by age, overall and by group, per country
plot_trajectory_sex <- function(df, ctry) {
  # split version
  df |>
    filter(!is.na(cladder), !is.na(equivalent_age)) |>
    mutate(chsex = as_factor(chsex)) |> 
    group_by(equivalent_age, chsex) |>
    summarise(mean_cl = mean(cladder), n = n(), .groups = "drop") |>
    ggplot(aes(equivalent_age, mean_cl, colour = chsex)) +
    geom_line() + geom_point() +
    scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
    scale_y_continuous(limits = c(3, 7)) +
    labs(title = ctry, x = "Age", y = "Mean life satisfaction") +
    theme_minimal()
}

plots <- imap(stacked, plot_trajectory_sex)
wrap_plots(plots, ncol = 2)
```

*By majority/minority status and sex*

```{r traj-sex-minority}
#| code-fold: true

bind_rows(stacked, .id = "country") |>
  filter(!is.na(cladder), !is.na(equivalent_age)) |>
  mutate(chsex = as_factor(chsex)) |>
  group_by(country, equivalent_age, chsex, minority) |>
  summarise(mean_cl = mean(cladder),
            se = sd(cladder) / sqrt(n()),
            .groups = "drop") |>
  ggplot(aes(equivalent_age, mean_cl, colour = minority, fill = minority)) +
  geom_ribbon(aes(ymin = mean_cl - 1.96 * se, ymax = mean_cl + 1.96 * se),
              alpha = 0.2, colour = NA) +
  geom_line() + geom_point() +
  scale_colour_manual(values = group_cols) +
  scale_fill_manual(values = group_cols) +
  facet_grid(country ~ chsex) +
  scale_x_continuous(breaks = c(8, 12, 15, 19, 22, 26, 29)) +
  scale_y_continuous(limits = c(4, 7)) +
  theme_minimal()
```

------------------------------------------------------------------------

### Psychological distress (SDQ-E) - 1994-cohort only

We cannot visualise SDQ properly because the measure is not invariant across time so descriptive visualisations would not be comparable.

We can only do per-item plots, for each round across countries. In Peru the same scale is used across countries so it can be plotted.

```{r sdq-country}
# Long format: one row per person-item-wave, with the raw response category
item_long <- imap_dfr(stacked, function(df, ctry) {
  df |>
    filter(yc == 0) |>
    select(childid, any_of(as.vector(outer(sdq_items, c("r3","r4","r5"),
                                            \(a,b) paste0(a,"_",b))))) |>
    pivot_longer(-childid, names_to = "item_wave", values_to = "response") |>
    filter(!is.na(response)) |>
    mutate(country = ctry,
           item = str_remove(item_wave, "_r[3-5]$"),
           wave = dplyr::recode(str_extract(item_wave, "r[3-5]$"),
                                r3 = "Age 15", r4 = "Age 19", r5 = "Age 22"))
})

# Proportion in each response category, per country x item x wave
item_props <- item_long |>
  count(country, item, wave, response) |>
  group_by(country, item, wave) |>
  mutate(prop = n / sum(n)) |>
  ungroup()

# Plot one item at a time (e.g. worry), facet by wave, countries side by side
item_props |>
  filter(item == "worry") |>
  ggplot(aes(x = factor(response), y = prop, fill = wave)) +
  geom_col(position = "dodge") +
  facet_wrap(~ country) +
  labs(title = "SDQ item: worry — response distribution across waves, by country",
       x = "Response category", y = "Proportion", fill = "Wave") +
  scale_fill_manual(values = c("Age 15" = "#4477AA",
                               "Age 19" = "#CC6677",
                               "Age 22" = "#DDCC77")) +
  theme_minimal()

# Peru only
# Peru only: mean response per item across waves (scale-consistent, so comparable)
item_long |>
  filter(country == "peru") |>
  group_by(item, wave) |>
  summarise(mean_resp = mean(response), se = sd(response)/sqrt(n()), .groups = "drop") |>
  ggplot(aes(x = wave, y = mean_resp, group = item, colour = item)) +
  geom_line() + geom_point() +
  geom_errorbar(aes(ymin = mean_resp - 1.96*se, ymax = mean_resp + 1.96*se), width = 0.1) +
  labs(title = "Peru: mean SDQ item response across waves (3-point scale throughout)",
       x = NULL, y = "Mean response (1–3)", colour = "Item") +
  theme_minimal()
```

```{r sdq-maj-min}

# Check size of smallest response category by majority/minority status - are some categories very sparse?
item_long_grp |>
  count(country, wave, item, response) |>          # NOT split by group
  group_by(country, wave, item) |>
  summarise(min_cell = min(n), .groups = "drop") |>
  arrange(min_cell) |> head(20)

# Raw counts per response category, split by group — look for tiny cells
sparse_check <- item_long_grp |>
  count(country, group, wave, item, response) |>
  tidyr::pivot_wider(names_from = response, values_from = n, values_fill = 0)

# Flag the worry item as an example; scan all
sparse_check |> dplyr::filter(item == "worry") |> print(n = Inf)

# Quickest summary: smallest cell per country x group x wave x item
item_long_grp |>
  count(country, group, wave, item, response) |>
  group_by(country, group, wave, item) |>
  summarise(min_cell = min(n), .groups = "drop") |>
  arrange(min_cell) |>
  head(20)
```

Looks alright, even for the smallest bin there are 90 people, and enough of individuals from minority group.

------------------------------------------------------------------------

## Descriptive statistics

*Summary table: per country x age x religious majority/minority*

```{r table-descriptives}
#| code-fold: true

# All countries together
all_df <- bind_rows(stacked, .id = "country")

# Pivot longer, mean/SD/n per country x age x minority group
desc <- all_df |> 
  pivot_longer(c(cladder, sdq), names_to = "outcome", values_to = "value") |> 
  filter(!is.na(value), !is.na(minority)) |> 
  group_by(country, outcome, equivalent_age, minority) |> 
  summarise(mean = mean(value), sd = sd(value), n = n(), .groups = "drop")

# Clear wide table - maj/min side by side
desc_table <- desc |> 
  mutate(cell = sprintf("%.2f (%.2f) [n=%d]", mean, sd, n)) |> 
  select(country, outcome, equivalent_age, minority, cell) |> 
  pivot_wider(names_from = minority, values_from = cell) |> 
  arrange(country, outcome, equivalent_age)

desc_table
```

*Group comparisons at each timepoint - Welch's t-test*

```{r comparisons}
#| code-fold: true

comparisons <- all_df |> 
  pivot_longer(c(cladder, sdq), names_to = "outcome", values_to = "value") |> 
  filter(!is.na(value), !is.na(minority)) |> 
  mutate(minority = factor(minority, levels = c("Minority", "Majority"))) |> 
  group_by(country, outcome, equivalent_age) |> 
  filter(n_distinct(minority) == 2,
         min(table(minority)) >= 2) |> 
  summarise(
    n_min = sum(minority == "Minority"),
    n_maj = sum(minority == "Majority"),
    tt = list(t.test(value ~ minority)),
    .groups = "drop"
  ) |> 
  mutate(
    diff_min_maj = map_dbl(tt, ~ .x$estimate[1] - .x$estimate[2]),
    ci_low = map_dbl(tt, ~ .x$conf.int[1]),
    ci_high = map_dbl(tt, ~ .x$conf.int[2]),
    p_value = map_dbl(tt, "p.value")
  ) |> 
  select(-tt)
comparisons
```

Welch's t test is preferred since the minority and majority groups do not have equal variances nor sample sizes. We can test whether the group means between minorities and majorities are statistically significant.

Note that these are purely descriptive, we're doing many independent t-tests with no correction so this is more for establishing overall patterns.

*diff_min_maj* is the minority - majority, i.e. if *diff_min_maj* is positive for *cladder*, that means that minorities have higher life satisfaction; whereas if it's positive for *sdq*, that means minorities have lower psychological distress (since the two scales are coded differently).

Summary

- Ethiopia: cladder at age 15, 22, 26, 29: minority better life satisfaction than majority

- India: cladder at age 22: minority better life satisfaction than majority

------------------------------------------------------------------------

## Variability

```{r histograms}
#| code-fold: true

plot_hist <- function(df, ctry, outcome_var) {
  df |> 
    filter(!is.na(.data[[outcome_var]]), !is.na(minority)) |> 
    ggplot(aes(x = .data[[outcome_var]], fill = minority)) +
    geom_histogram(aes(y = after_stat(density)),
                   binwidth = 1, alpha = 0.5, position = "identity") +
    facet_wrap(~ equivalent_age, ncol = 4) +
    labs(title = paste0(ctry, " - ", outcome_var), x = outcome_var, y = "Count") +
    theme_minimal() +
    scale_fill_manual(values = group_cols) +
    theme(legend.position = "none")
}

# Life satisfaction (combined cohort: 7 ages x 2 groups)
iwalk(stacked, ~ print(plot_hist(.x, .y, "cladder")))

# Psychological distress (1994-cohort: 3 ages x 2 groups)
iwalk(stacked, ~ print(plot_hist(.x, .y, "sdq")))
```

------------------------------------------------------------------------

------------------------------------------------------------------------

# Latent Growth Curve Analyses (LGCA)

*Structure data*

```{r pivot-wide}
#| code-fold: true

sdq_items <- c("headache","worry","unhappy","nervous","fears")
item_cols <- as.vector(outer(sdq_items, c("r3","r4","r5"), \(a,b) paste0(a,"_",b)))

to_wide <- function(df) {
  wide_scores <- df |>
    mutate(chsex = as.numeric(as_factor(chsex)) - 1,  # 0/1
           minority = if_else(minority == "Minority", 1, 0)) |>
    select(childid, chsex, minority, clustid, yc, equivalent_age, cladder) |>
    pivot_wider(names_from = equivalent_age,
                values_from = cladder,
                names_prefix = "cladder_age")
  wide_items <- df |>  # specific sdq items per round
    select(childid, any_of(item_cols)) |> 
    group_by(childid) |> 
    summarise(across(any_of(item_cols), ~first(na.omit(.x))), .groups = "drop")
  left_join(wide_scores, wide_items, by = "childid")
}
wide <- map(stacked, to_wide)
wide <- map(wide, ~ .x |> group_by(childid) |>
              fill(clustid, .direction = "downup") |> ungroup())

names(wide$ethiopia)  # check column names
```

*Filter children* (only those with 3 complete cladder out of 5 for 2001-cohort, 7 for 1994-cohort)

```{r filter-sample}
#| code-fold: true

wide <- map(wide, ~ .x |> 
              mutate(n_cl = rowSums(!is.na(across(starts_with("cladder_age"))))) |> 
              filter(n_cl >= 3) |> 
              select(-n_cl))
```

## Life satisfaction models

Given the two cohorts are not equivalent - with diverging trajectories, we will fit the models separately, and religious minority status and sex predicting the growth factors. 2001-cohort will have age 8-22, and 1994-cohort 12-29.

We have two cohorts across four countries, so for simplicity we will build a function so it can be easily replicated.

*LGCA helpers*

```{r lgca-helpers}
#| code-fold: true

# Build lavaan growth syntax from a vector of ages, centred at the first age.
# type: "intercept", "linear", or "conditional"
make_spec <- function(ages, shape = "linear", conditional = FALSE, extra = NULL) {
  t    <- (ages - ages[1]) / 10          # decades 
  cols <- paste0("cladder_age", ages)
  i    <- paste0("i =~ ", paste0("1*", cols, collapse = " + "))

  if (shape == "intercept") {
    spec <- i; factors <- "i"

  } else if (shape == "basis") {                       # latent-basis (freed-loading)
    load <- rep("NA", length(ages))                    # middle loadings free
    load[1] <- "0"; load[length(ages)] <- "1"          # first fixed 0, last fixed 1
    s <- paste0("s =~ ", paste0(load, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")

  } else {
    s <- paste0("s =~ ", paste0(t, "*", cols, collapse = " + "))
    spec <- paste(i, s, sep = "\n"); factors <- c("i", "s")
    if (shape %in% c("quadratic", "cubic")) {
      q <- paste0("q =~ ", paste0(t^2, "*", cols, collapse = " + "))
      spec <- paste(spec, q, sep = "\n"); factors <- c(factors, "q")
    }
    if (shape == "cubic") {
      cub <- paste0("cub =~ ", paste0(t^3, "*", cols, collapse = " + "))
      spec <- paste(spec, cub, sep = "\n"); factors <- c(factors, "cub")
    }
  }

  if (conditional) {
    spec <- paste(c(spec, paste0(factors, " ~ minority + chsex")), collapse = "\n")
  }
  if(!is.null(extra)){
    spec <- paste(c(spec, extra), collapse ="\n")
  }
  spec
}

# Ages available per cohort per country
ages_comb <- list(ethiopia = c(8,12,15,19,22,26,29), india = c(8,12,15,19,22,26,29),
                peru = c(8,12,15,19,22,26,29), vietnam = c(8,12,15,19,22,26))
ages_yc <- list(ethiopia = c(8,12,15,19,22), india = c(8,12,15,19,22),
                peru = c(8,12,15,19,22), vietnam = c(8,12,15,19)) # vietnam up to age 19 only
ages_oc <- list(ethiopia = c(12,15,19,22,26,29), india = c(12,15,19,22,26,29),
                peru = c(12,15,19,22,26,29), vietnam = c(12,15,19,22,26)) # vietnam up to age 26 only

# Fit one model type for one cohort, across all countries
fit_all <- function(ages_list, shape, cohort_val, conditional = FALSE, constraints = NULL) {
  imap(wide, function(df, ctry) {
    spec <- make_spec(ages_list[[ctry]], shape, conditional = conditional,
                      extra = constraints[[ctry]])
    if (!identical(cohort_val, "all")) {
      df <- filter(df, yc == cohort_val)
    }
    d <- df |>
      filter(!is.na(minority), !is.na(chsex)) |>
      mutate(minority = as.numeric(minority), chsex = as.numeric(chsex))

    tryCatch(growth(spec, data = d, missing = "fiml"), error = function(e) NULL)
  })
}

get_fit <- function(fits) {
  extract <- function(f) {
    if (is.null(f)) return(NULL)
    if (!lavInspect(f, "converged")) return(NULL)
    as_tibble(as.list(fitMeasures(f, c(
      "chisq", "df", "pvalue",           # model chi-square
      "cfi", "tli",
      "rmsea", "rmsea.ci.lower", "rmsea.ci.upper",   # RMSEA + its 90% CI
      "srmr",                           
      "aic", "bic", "bic2"               # bic2 = sample-size-adjusted BIC
    ))))
  }
  map_dfr(fits, extract, .id = "country")
}

get_effects <- function(fits) {
  extract <- function(f) {
    if (is.null(f)) return(NULL)
    if (!lavInspect(f, "converged")) return(NULL)
    parameterEstimates(f) |>
      filter(rhs %in% c("minority", "chsex"),
             lhs %in% c("i", "s", "q", "cub")) |>  
      select(lhs, rhs, est, se, pvalue)
  }
  map_dfr(fits, extract, .id = "country")
}

check_model <- function(fit) {
  if (is.null(fit)) return("NULL (errored)")
  if (!lavInspect(fit, "converged")) return("did not converge")
  pe   <- parameterEstimates(fit)
  vars <- pe[pe$op == "~~" & pe$lhs == pe$rhs, ]     # all variances
  if (any(vars$est < 0)) return("Heywood (negative variance)")
  "OK"
}
```

### Combined cohort

*Intercept only*

```{r combined}
comb_lin   <- fit_all(ages_comb, "linear",    "all")
comb_quad  <- fit_all(ages_comb, "quadratic", "all")
comb_basis <- fit_all(ages_comb, "basis",     "all", constraints = list(india = "s ~~ 0*s")) # slope-factor variance was -0.002, need to constraint so model fits

fit_table <- function(fits, model) {
  tibble(country = names(fits)) |>
    mutate(status = purrr::map_chr(fits, check_model)) |>
    left_join(get_fit(fits), by = "country") |>
    mutate(model = model)
}

# Compare models
bind_rows(
  fit_table(comb_lin,   "Linear"),
  fit_table(comb_quad,  "Quadratic"),
  fit_table(comb_basis, "Basis")
) |>
  relocate(country, model, status) |>
  mutate(
    across(c(cfi, tli, srmr, rmsea, rmsea.ci.lower, rmsea.ci.upper), ~ round(.x, 3)),
    across(c(chisq, aic, bic, bic2), ~ round(.x, 1)),
    pvalue = round(pvalue, 3)
  ) |>
  arrange(model, factor(model, levels = c("Linear", "Quadratic", "Basis")))

# Conditional model
comb_cond <- fit_all(ages_comb, "basis", "all", conditional = TRUE,
                     constraints = list(india = "s ~~ 0*s"))  
get_effects(comb_cond)
```

### 2001-cohort age 8-22

*Intercept only*

```{r yc-intercept}
yc_int <- fit_all(ages_yc, "intercept", 1)
get_fit(yc_int)
```

*Linear (intercept and slope), FIML with clustering by clustid*

```{r yc-linear}
yc_lin <- fit_all(ages_yc, "linear", 1)
get_fit(yc_lin)
```

*Conditional model (minority and sex)*

```{r yc-conditional}
yc_cond <- fit_all(ages_yc, "linear", 1, conditional = TRUE)
get_effects(yc_cond)
```

*Quadratic*

```{r quadratic}
yc_quad <- fit_all(ages_yc, "quadratic", 1)
get_fit(yc_quad)

# No cubic for 2001-cohort; not enough time-points
```

*Conditional, quadratic*

```{r yc-conditional-quadratic}
yc_cond <- fit_all(ages_yc, "quadratic", 1, conditional = TRUE)
get_effects(yc_cond)
```

*Basis model*

```{r yc-basis}
yc_basis <- fit_all(ages_yc, "basis", 1,
                    constraints = list(ethiopia = "s ~~ 0*s")) # specify ethiopia constraints separately
purrr::map_chr(yc_basis, check_model)     
get_fit(yc_basis)

yc_basis_cond <- fit_all(ages_yc, "basis", 1, conditional = TRUE,
                   constraints = list(ethiopia = "s ~~ 0*s"))
get_effects(yc_basis_cond)
```

We use basis model / freed-loading model because trajectories were not monotonic - it seemed to decline throughout adolescence but partially recover over time. Linear and quadratic models fit poorly, but latent-basis (freed-loading) growth model can estimate the shape of average trajectory from the data rather than imposing a functional form.

In basis models, the first and last factor loadings are fixed to 0 and 1 respectively and in between loadings are freely estimated. Intercept = expected outcome at initial stage; slope = total change across observed age range; estimated loadings = proportion of total change realised at each save.

Model fit is now substantially improved with the freed-loading specification.

In Ethiopia 2001-cohort, the slope-factor variance was non-significant and marginally negative (Heywood case), suggesting negligible variability in trajectory of change between individauls. To fit a model we thus have to constrain the slope-factor variance to 0. This only affects estimation of individual differences in slope, not the mean trajectory or estimated minority-majority effects.

------------------------------------------------------------------------

### 1994-cohort age 12-29

*Intercept-only*

```{r oc-intercept}
oc_int <- fit_all(ages_oc, "intercept", 0)
get_fit(oc_int)
```

*Linear (intercept + slope)*

```{r oc-linear}
oc_lin <- fit_all(ages_oc, "linear", 0)
get_fit(oc_lin)
```

*Conditional (minority + sex)*

```{r oc-conditional}
oc_cond <- fit_all(ages_oc, "linear", 0, conditional = TRUE)
get_effects(oc_cond)
```

*Quadratic and cubic*

```{r quad-cub}
oc_quad <- fit_all(ages_oc, "quadratic", 0)
oc_cubic <- fit_all(ages_oc, "cubic", 0)   

get_fit(oc_quad)
get_fit(oc_cubic)

# It seems cubic failed to converge for all countries
```

*Conditional, quadratic*

```{r oc-conditional}
oc_cond <- fit_all(ages_oc, "quadratic", 0, conditional = TRUE)
get_effects(oc_cond)
```

*Basis model*

```{r oc-basis}
oc_basis <- fit_all(ages_oc, "basis", 0) 
purrr::map_chr(oc_basis, check_model) 
get_fit(oc_basis)

oc_basis_cond <- fit_all(ages_oc, "basis", 0, conditional = TRUE)
get_effects(oc_basis_cond)
```

### Check fit indices

```{r fit-indices}
bind_rows(
  get_fit(yc_lin)  |> mutate(cohort = "Younger", model = "Linear"),
  get_fit(oc_lin)  |> mutate(cohort = "Older",   model = "Linear"),
  get_fit(yc_quad) |> mutate(cohort = "Younger", model = "Quadratic"),
  get_fit(oc_quad) |> mutate(cohort = "Older",   model = "Quadratic"),
  get_fit(yc_basis) |> mutate(cohort = "Younger", model = "Basis"),
  get_fit(oc_basis) |> mutate(cohort = "Older", model = "Basis")
) |>
  relocate(country, cohort, model) |>
  mutate(across(c(cfi, tli, rmsea), ~ round(.x, 3)))
```

Fit indices are... poor. CFI and TLI are not \>=0.95, RMSEA mostly not \<=0.06 (except Vietnam).

Even for linear the fit is poor, and quadratic does not do any better. Basis better?

------------------------------------------------------------------------

### Trajectory visualisations

Visualise model-implied trajectories of minority vs majority groups per country, using basis model.

```{r traj-basis}
implied_traj_basis <- function(fit, ages) {
  if(is.null(fit)) return(NULL)
  if(!lavInspect(fit, "converged")) return(NULL)
  
  est <- parameterEstimates(fit)
  cols <- paste0("cladder_age", ages)
  
  # estimated basis loadings on s (freed loadings), pulled in wave order
  load <- sapply(cols, function(cc)
    est$est[est$op == "=~" & est$lhs == "s" & est$rhs == cc])
  
  cf <- coef(fit)
  maj_i <- cf["i~1"]; maj_s <- cf["s~1"]
  min_i <- maj_i + cf["i~minority"]
  min_s <- maj_s + cf["s~minority"]
  
  tibble(
    age = rep(ages, 2),
    group = rep(c("Majority", "Minority"), each = length(ages)),
    fit = c(maj_i + load * maj_s, # majority: intercept + loading x slope
            min_i + load * min_s) # minority: same, shifted by minority coefs
  )
}
```

```{r plot-basis}
plot_obs_model <- function(fits, ages_list, cohort_val, title) {
  model_df <- imap_dfr(fits, function(f, ctry) {
    d <- implied_traj_basis(f, ages_list[[ctry]])
    if (is.null(d)) return(NULL)
    mutate(d, country = ctry)
  })

  obs_df <- imap_dfr(stacked, function(df, ctry) {
    df |> filter(yc == cohort_val, !is.na(cladder),
                 equivalent_age %in% ages_list[[ctry]]) |>
      mutate(group = if_else(minority == "Minority", "Minority", "Majority")) |>
      group_by(country = ctry, age = equivalent_age, group) |>
      summarise(mean_cl = mean(cladder),
                se = sd(cladder) / sqrt(n()),          # <- SE of the mean
                .groups = "drop") |>
      mutate(lo = mean_cl - 1.96 * se,
             hi = mean_cl + 1.96 * se)
  })

  ggplot() +
    geom_ribbon(data = obs_df,
                aes(age, ymin = lo, ymax = hi, fill = group),
                alpha = 0.15, colour = NA) +               # <- the bands
    geom_point(data = obs_df, aes(age, mean_cl, colour = group), alpha = 0.6) +
    geom_line(data = model_df, aes(age, fit, colour = group), linewidth = 1) +
    facet_wrap(~ country) +
    scale_colour_manual(values = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
    scale_fill_manual(values   = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
    labs(title = title, y = "Life satisfaction", x = "Age", colour = NULL, fill = NULL) +
    theme_minimal()
}

plot_obs_model(yc_basis_cond, ages_yc, 1, "2001-Cohort")
plot_obs_model(oc_basis_cond, ages_oc, 0, "1994-Cohort")
```

------------------------------------------------------------------------

## Psychological distress models

With SDQ we will do both a measurement and structural model together, because the response options are different across ages within the same country, we will construct a latent variable emotional symptoms using the five SDQ items for each timepoint, and moel that.

WLSMV (Weighted Least Squares Mean and Variance) - for CFA here; SDQ-E is ordinal.

### *Measurement model*

*Create latent variable SDQ-E using 5 individual items* Per wave per country

```{r measurement-model-sdq}
#| code-fold: true

# Confirmatory factor analysis
fit_cfa_wave <- function(df, r) {
  items <- paste0(sdq_items, "_", r)
  spec  <- paste0("emo =~ ", paste(items, collapse = " + "))
  d <- df |> filter(yc == 0)                    # older cohort only
  tryCatch(
    lavaan::cfa(spec, data = d, ordered = items,
                estimator = "WLSMV", missing = "pairwise"),
    error = function(e) NULL)
}

# All countries for all 3 waves
cfa_singlewave <- imap(wide, function(df, ctry) {
  list(age15 = fit_cfa_wave(df, "r3"),
       age19 = fit_cfa_wave(df, "r4"),
       age22 = fit_cfa_wave(df, "r5"))
})
purrr::map(cfa_singlewave, ~ purrr::map_chr(.x, check_model))

# Check fit and loadings
purrr::imap(cfa_singlewave, function(waves, ctry) {
  purrr::map_dfr(waves, function(f) as_tibble(as.list(fitMeasures(f, c("cfi","tli","rmsea","srmr")))),
                 .id = "wave") |> mutate(country = ctry)
})

# Example: Ethiopia age 15 — should all be positive, ideally >.5
standardizedSolution(cfa_singlewave$ethiopia$age15) |>
  dplyr::filter(op == "=~") |> dplyr::select(rhs, est.std, pvalue)
```

*Test longitudinal invariance*

```{r sdq-invariance}
#| code-fold: true

# 3-wave longitudinal CFA at a given invariance level 
# level: "configural" (nothing constrained), "metric" (loadings equal), "scalar "loadings + thresholds equal)
# n_thresh = number of thresholds a country's items have at waves 19/22 - Peru is 3pt i.e. 2 thresholds, remaining are 4pt i.e. 3 thresholds

make_long_cfa <- function(level = "configural", n_thresh = 3) {
  waves <- c("15" = "r3", "19" = "r4", "22" = "r5")
  facs <- paste0("e", names(waves)) # e15, e19, e22
  lines <- character(0)
  
  # measurement model per wave
  for (k in seq_along(waves)) {
    it <- paste0(sdq_items, "_", waves[k])
    labs <- if(level == "configural") paste0("L", seq_along(sdq_items), "_w", k)
            else                      paste0("L", seq_along(sdq_items)) 
    lines <- c(lines, paste0(facs[k], " =~ ", paste0(labs, "*", it, collapse = " + ")))
  }
  
  # full scalar
  if(level == "scalar") {
    for (itm in sdq_items) {
      # constrain shared thresholds across three waves
      i15 <- paste0(itm, "_r3"); i19 <- paste0(itm, "_r4"); i22 <- paste0(itm, "_r5")
      lines <- c(lines,
                 paste0(i15, " | th1*t1"), paste0(i19, " | th1*t1"), paste0(i22, " | th1*t1"),
                 paste0(i15, " | th2*t2"), paste0(i19, " | th2*t2"), paste0(i22, " | th2*t2"))
    }
  }
  if(level == "partial") {
    # partial scalar, free age 15
    for (itm in sdq_items) {
      i19 <- paste0(itm,"_r4"); i22 <- paste0(itm,"_r5")
      for (th in seq_len(n_thresh)) { # 2 for Peru, 3 for others
        lbl <- paste0("th", itm, th)
        lines <- c(lines,
                 paste0(i19, " | ", lbl, "*t", th), 
                 paste0(i22, " | ", lbl, "*t", th)) 
                 # age 15 not listed so can be freely estimated
    }
    }
  }    
  
  # correlated residuals
  for (itm in sdq_items) {
    cols <- paste0(itm, "_", waves)
    lines <- c(lines,
               paste0(cols[1], " ~~ ", cols[2]),
               paste0(cols[1], " ~~ ", cols[3]),
               paste0(cols[2], " ~~ ", cols[3]))
  }
  paste(lines, collapse = "\n")
}

# Fit all four levels for one country; n_thresh set per country
fit_invariance <- function(df, n_thresh = 3) {
  items <- as.vector(outer(sdq_items, c("r3", "r4", "r5"), \(a,b) paste0(a,"_",b)))
  d <- df |> filter(yc==0)
  fit_one <- function(level)
    tryCatch(
    lavaan::cfa(make_long_cfa(level, n_thresh = n_thresh), data = d, ordered = items,
                estimator = "WLSMV", missing = "pairwise",
                parameterization = "theta"),
    error = function(e) NULL)
  list(
    configural = fit_one("configural"),
    metric = fit_one("metric"),
    scalar = fit_one("scalar"),
    partial = fit_one("partial"))
}

# per country threshold count
n_thresh_by_country <- c(ethiopia = 3, india = 3, peru = 2, vietnam = 3)
```

The drop from configural -\> metric is \~0.01 so that's okay. Configural and metric invariance hold - SDQ-E measures the same construct across the three waves.

However, from metric -\> scalar the change is 0.20 i.e. fail; this is expected because of the change in threshold for three countries, from a 3pt to 4pt response. Age 15 SDQ is not comparable to 19/22.

The only solution is to adopt partial scalar invariance, meaning we constrain the thresholds across 19-22 (same scale) but freely estimate the item thresholds at age 15. This retains all three waves while accounting for measurement change. However we still need to interpret the age 15 latent factor with appropriate caution.

For Ethiopia, India, and Vietnam, it makes sense to have partial invariance structure because of change in threshold across rounds. Interesting that despite uniform scaling across time in Peru full scalar did not hold; for consistency across countries we free age-15 for all, including Peru. Though for Peru the reasons are different - perhaps a developmental (age-related) threshold shift where interpretation of the scale differs across ages.

*Partial scalar invariance*

```{r partial-scalar-invariance}

inv_eth <- fit_invariance(wide$ethiopia, n_thresh = n_thresh_by_country["ethiopia"])
inv_ind <- fit_invariance(wide$india,    n_thresh = n_thresh_by_country["india"])
inv_per <- fit_invariance(wide$peru,     n_thresh = n_thresh_by_country["peru"])
inv_vnm <- fit_invariance(wide$vietnam,  n_thresh = n_thresh_by_country["vietnam"])

# admissibility + the full configural -> metric -> scalar -> partial table per country
list(ethiopia = inv_eth, india = inv_ind, peru = inv_per, vietnam = inv_vnm) |>
  purrr::imap_dfr(function(inv, ctry)
    compare_invariance(inv) |> mutate(country = ctry))
```

Partial invariance recovered well for all which is good news!

### *Growth model*

Second-order growth model built on partial-invariance measurement structure

```{r growth-model}
# Growth: intercept + linear slope on three wave-factors
make_spec_sdq <- function(shape = "linear", conditional = FALSE,
                          extra = NULL, n_thresh = 3) {
  waves <- c("15" = "r3", "19" = "r4", "22" = "r5")
  ages <- c(15, 19, 22)
  facs <- paste0("e", names(waves))
  t <- (ages - ages[1]) / 10
  lines <- character(0)
  
  # measurement: metric loadings
  for (k in seq_along(waves)) {
    it <-  paste0(sdq_items, "_", waves[k])
    labs <- paste0("L", seq_along(sdq_items))
    lines <- c(lines, paste0(facs[k], " =~ ", paste0(labs, "*", it, collapse = " + ")))
  }
  
  # partial scalar thresholds
  for (itm in sdq_items) {
    i19 <- paste0(itm, "_r4"); i22 <- paste0(itm, "_r5")
    for (th in seq_len(n_thresh)) {
      lbl <- paste0("th", itm, th)
      lines <- c(lines, paste0(i19, " | ", lbl, "*t", th),
                        paste0(i22, " | ", lbl, "*t", th))
    }
  }
  
  # correlated residuals
  for (itm in sdq_items) {
    cols <- paste0(itm, "_", waves)
    lines <- c(lines, paste0(cols[1], " ~~ ", cols[2]),
               paste0(cols[1], " ~~ ", cols[3]),
               paste0(cols[2], " ~~ ", cols[3]))
  }
  
  # growth factors on three-wave factors
  i <- paste0("i =~ ", paste0("1*", facs, collapse = " + "))
  if (shape == "intercept") { g <- i; gf <- "i" }
  else {  # linear - only 3 waves
    s <- paste0("s =~ ", paste0(t, "*", facs, collapse = " + "))
    g <- paste(i, s, sep = "\n"); gf <- c("i", "s")
  }
  lines <- c(lines, g)
  
  if (conditional) lines <- c(lines, paste0(gf, " ~ minority + chsex"))
  if(!is.null(extra)) lines <- c(lines, extra)
  paste(lines, collapse = "\n")
}

n_thresh_by_country <- c(ethiopia = 3, india = 3, peru = 2, vietnam = 3)

fit_sdq <- function(shape, conditional = FALSE, constraints = NULL) {
  items <- as.vector(outer(sdq_items, c("r3","r4","r5"), \(a,b) paste0(a,"_",b)))
  imap(wide, function(df, ctry) {                    # loops over 4 countries
    d <- df |> filter(yc == 0, !is.na(minority), !is.na(chsex)) |>
      mutate(minority = as.numeric(minority), chsex = as.numeric(chsex))
    spec <- make_spec_sdq(shape, conditional, extra = constraints[[ctry]],
                           n_thresh = n_thresh_by_country[[ctry]])
    tryCatch(
      lavaan::growth(spec, data = d, ordered = items, estimator = "WLSMV",
                     missing = "pairwise", parameterization = "theta"),
      error = function(e) NULL)
  })
}
```

### Structural model

*Intercept-only and linear*

```{r int-lin-sdq}
sdq_int <- fit_sdq("intercept")
sdq_lin <- fit_sdq("linear", constraints = list(
  ethiopia = "i ~~ 0*i", # intercept variance going negative, need to constrain
  india = "s ~~ 0*s")) # slope variance constrained

purrr::map_chr(sdq_int, check_model)
purrr::map_chr(sdq_lin, check_model)
get_fit(sdq_lin)
```

In Ethiopia, intercept variance was non-significant and marginally negative (Heywood case), so we constrained to 0 - indicating negligible estimable individual variability in baseline SDQ-E. Avoided inadmissible solution. In India, slope variance was similarly constrained, reflecting the limited capacity of a 3-wave model to estimate individual differences in growth factors.

Given we only have 3 waves, a higher-order or freed-loading trajectory shape was not estimable. We specify only a linear slope, and the linear model fits well according to indices - CFI 0.96-0.99, RMSEA \<=0.043, SRMR \~0.05.

*Conditional (minority + sex)*

```{r conditional-sdq}
sdq_cond <- fit_sdq("linear", conditional = TRUE, constraints = list(
  ethiopia = "i ~~ 0*i",
  india    = "s ~~ 0*s"))
purrr::map_chr(sdq_cond, check_model)  
get_effects(sdq_cond)
```

### Trajectory visualisations

*Plot latent factor trajectory*

SDQ is not a latent factor and not observed score, which means we are plotting the model-implied latent factor means at each wave.

**By religious minority/majority**

```{r traj-sdq-rel}
implied_traj_sdq <- function(fit, ages = c(15,19,22)) {
  if (is.null(fit) || !lavInspect(fit, "converged")) return(NULL)
  cf <- coef(fit)
  t <- (ages - ages[1]) / 10 # 0, 0.4, 0.7 - matches slope coding
  
  # intercept and slope means for reference
  maj_i <- cf[["i~1"]]; maj_s <- cf[["s~1"]]
  # minority (holding sex=0)
  min_i <- maj_i + cf[["i~minority"]]
  min_s <- maj_s + cf[["s~minority"]]
  
  tibble(
    age = rep(ages,2),
    group = rep(c("Majority", "Minority"), each = length(ages)),
    fit = c(maj_i + maj_s * t, min_i + min_s * t)
  )
}

sdq_model_df <- imap_dfr(sdq_cond, function(f, ctry) {
  d <- implied_traj_sdq(f)
  if (is.null(d)) return(NULL)
  mutate(d, country = ctry)
})

sdq_annot <- get_effects(sdq_cond) |>
  dplyr::filter(lhs == "s", rhs == "minority") |>
  dplyr::mutate(label = paste0("Minority slope p = ", sprintf("%.2f", pvalue)))

ggplot(sdq_model_df, aes(age, fit, colour = group)) +
  geom_line(linewidth = 1) + 
  geom_point() +
  geom_text(data = sdq_annot, aes(x = 15, y = Inf, label = label), colour = "grey30", hjust = 0, vjust = 1.4, size = 3, inherit.aes = FALSE) +
  facet_wrap(~ country) +
  scale_colour_manual(values = c("Majority" = "#4477AA", "Minority" = "#CC6677")) +
  scale_x_continuous(breaks = c(15, 19, 22)) +
  labs(title = "Psychological distress (SDQ-E) model-implied latent trajectories",
       subtitle = "Religious minority vs majority latent trajectories",
       caption = "Higher values indicate greater distress. Estimates adjust for sex. Trajectories are model-implied estimates.",
       y = "Latent psychological distress (model-implied)", x = "Age", colour = NULL) +
  theme_minimal()
```

**By sex**

```{r traj-sdq-sex}
implied_traj_sex <- function(fit, ages = c(15, 19, 22)) {
  if (is.null(fit) || !lavInspect(fit, "converged")) return(NULL)
  cf <- coef(fit)
  t <- (ages - ages[1]) / 10
  
  male_i <- cf[["i~1"]]; male_s <- cf[["s~1"]] # reference: male, religous majority
  # minority (holding minority=0)
  fem_i <- male_i + cf[["i~chsex"]]
  fem_s <- male_s + cf[["s~chsex"]]
  
  tibble(
    age = rep(ages,2),
    group = rep(c("Male", "Female"), each = length(ages)),
    fit = c(male_i + male_s * t, fem_i + fem_s * t)
  )
}

sdq_sex_df <- imap_dfr(sdq_cond, function(f, ctry) {
  d <- implied_traj_sex(f)
  if (is.null(d)) return(NULL)
  mutate(d, country = ctry)
})

sdq_sex_annot <- get_effects(sdq_cond) |> 
  dplyr::filter(lhs == "s", rhs == "chsex") |> 
  dplyr::mutate(label = paste0("Slope p = ", sprintf("%.3f", pvalue)))

ggplot(sdq_sex_df, aes(age, fit, colour = group)) +
  geom_line(linewidth = 1) + 
  geom_point() +
  geom_text(data = sdq_sex_annot, aes(x = 15, y = Inf, label = label), colour = "grey30", hjust = 0, vjust = 1.4, size = 3, inherit.aes = FALSE) +
  facet_wrap(~ country) +
  scale_colour_manual(values = c("Male" = "#009E73", "Female" = "#E69F00")) +
  scale_x_continuous(breaks = c(15, 19, 22)) +
  labs(title = "Psychological distress (SDQ-E) model-implied latent trajectories",
       caption = "Higher values indicate greater distress. Estimates adjust for religious minority status.",
       y = "Latent psychological distress (model-implied)", x = "Age", colour = NULL) +
  theme_minimal()
```

- **Ethiopia**: similar distress at age 15 (intercept b=-0.10, p=.065), but girls' distress increased significantly faster than boys (slope b=0.55, p\<.001).

- **India**: girls higher distress at 15 (intercept b=0.66, p\<.001) with no difference in change over time.

- **Peru**: boys higher distress at 15 (intercept b=-0.46, -\<.001) and worsened more quickly than girls over time (slope b=-0.48, p=.002).

- **Vietnam**: girls higher distress at 15 (intercept b=0.26, p\<.001) with no difference over time.
